### 2021-08-04 main/main.Rmd
init paper with some outlines

---
title: A template for the *arxiv* style
authors:
  - name: H.Sherry Zhang
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    department: Institute of Statistics
    affiliation: University of Natural Resources and Life Sciences
    location: Vienna, Austria
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    department: 34 Village Street, Docklands VIC 3008 Australia
    affiliation: CSIRO Data61 
    location: Melbourne, Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  patricia.menendez@monash.edu 
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction

Spatio-temporal data

Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis

sections

# Literature Review

Existing R pkgs for spatio/ temporal data

- `sp` + `spacetime` (time series data from package `xts`)

- `sf` 
  - simple features geometry 
  - only spatio
  - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure

- `stars`
  - a newer structure allow both spatio + temporal 
  - array

# A new data structure for spatio-temporal data

Intro to cubble:

- list-column: rowwise_df with temporal variables, including the time index, nested. 
  - Focus on spatio: those output per station

- long form: grouped_df 
  - Focus on temporal

Compatible with tidyverse manipulation and tsibble

# Examples

Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations across Australia

An exploratory data analysis questions: What's the climate profile look like in Australia

  - General features: Any general trend/ fluctuation in prcp, tmax, and tmin?
  - Local features: Any station stands out from the crowd?

## Manipulation

  - data quality check: filter out stations have variables not properly recorded 
  - data summary: 
    - daily -> monthly/ weekly, 
    - summarise by mean for tmax/ tmin, sum for prcp
  -

## Graphics

Static + interactive -> tooltip to show additional information upon hovering

  - Where are those stations on the map? 
    - Mention mostly aero, airport, and lighthouse

# Summary