### 2021-08-04 main/main.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{A template for the \emph{arxiv} style}

\author{
    H.Sherry Zhang
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}} \\
   \And
    Dianne Cook
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
   \And
    Ursula Laa
   \\
    Institute of Statistics \\
    University of Natural Resources and Life Sciences \\
  Vienna, Austria \\
  \texttt{\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}} \\
   \And
    Nicolas Langrené
   \\
    34 Village Street, Docklands VIC 3008 Australia \\
    CSIRO Data61 \\
  Melbourne, Australia \\
  \texttt{\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}} \\
   \And
    Patricia Menéndez
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}} \\
  }

% Pandoc citation processing

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
Enter the text of your abstract here.
\end{abstract}

\keywords{
    blah
   \and
    blee
   \and
    bloo
   \and
    these are optional and can be removed
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Spatio-temporal data

Motivation: Many focus on spatio/ temporal solely -\textgreater{}
spatio-temporal vector data structure for data analysis

sections

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

Existing R pkgs for spatio/ temporal data

\begin{itemize}
\item
  \texttt{sp} + \texttt{spacetime} (time series data from package
  \texttt{xts})
\item
  \texttt{sf}

  \begin{itemize}
  \tightlist
  \item
    simple features geometry
  \item
    only spatio
  \item
    good for geographical but these geometry types i.e.~multipolygon,
    can be hard to manipulate. Would be nice to have a vector data
    structure
  \end{itemize}
\item
  \texttt{stars}

  \begin{itemize}
  \tightlist
  \item
    a newer structure allow both spatio + temporal
  \item
    array
  \end{itemize}
\end{itemize}

\hypertarget{a-new-data-structure-for-spatio-temporal-data}{%
\section{A new data structure for spatio-temporal
data}\label{a-new-data-structure-for-spatio-temporal-data}}

Intro to cubble:

\begin{itemize}
\tightlist
\item
  list-column: rowwise\_df with temporal variables, including the time
  index, nested.

  \begin{itemize}
  \tightlist
  \item
    Focus on spatio: those output per station
  \end{itemize}
\item
  long form: grouped\_df

  \begin{itemize}
  \tightlist
  \item
    Focus on temporal
  \end{itemize}
\end{itemize}

Compatible with tidyverse manipulation and tsibble

\hypertarget{examples}{%
\section{Examples}\label{examples}}

Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations
across Australia

An exploratory data analysis questions: What's the climate profile look
like in Australia

\begin{itemize}
\tightlist
\item
  General features: Any general trend/ fluctuation in prcp, tmax, and
  tmin?
\item
  Local features: Any station stands out from the crowd?
\end{itemize}

\hypertarget{manipulation}{%
\subsection{Manipulation}\label{manipulation}}

\begin{itemize}
\tightlist
\item
  data quality check: filter out stations have variables not properly
  recorded
\item
  data summary:

  \begin{itemize}
  \tightlist
  \item
    daily -\textgreater{} monthly/ weekly,
  \item
    summarise by mean for tmax/ tmin, sum for prcp
  \end{itemize}
\item
\end{itemize}

\hypertarget{graphics}{%
\subsection{Graphics}\label{graphics}}

Static + interactive -\textgreater{} tooltip to show additional
information upon hovering

\begin{itemize}
\tightlist
\item
  Where are those stations on the map?

  \begin{itemize}
  \tightlist
  \item
    Mention mostly aero, airport, and lighthouse
  \end{itemize}
\end{itemize}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}