### 2021-08-04 main/main.Rmd
draft on existing spatio + temporal R pkgs

---
title: A template for the *arxiv* style
authors:
  - name: H.Sherry Zhang
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    department: Institute of Statistics
    affiliation: University of Natural Resources and Life Sciences
    location: Vienna, Austria
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    department: 34 Village Street, Docklands VIC 3008 Australia
    affiliation: CSIRO Data61 
    location: Melbourne, Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  patricia.menendez@monash.edu 
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction

Spatio-temporal data

<!-- While satellite imagery can record information on a continuous spatial domain, other spatial information are recorded at a fixed set of locations characterised by latitude, longitude, and other spatial metadata. Examples of this type of data includes climate measures from weather observation stations and river level data recorded by electronic gauges. Analysing this data requires less considerations on the geometry type and map projection while more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations. -->

Motivation: Many focus on spatio/ temporal solely -> spatio-temporal vector data structure for data analysis

sections

# Existing data structure for spatio and temporal data

There has been a large class of implementations dedicated to processing the spatial data. This includes `sf` [@pebesma2018simple] and its precedent `sp` [@pebesma2005s] for ... and `raster` [@raster] and `terra` [@terra] for raster data. While these implementations specialised in geographic manipulations with different type of simple features, it doesn't incorporate a temporal dimension in the data structure. Project like `stars` [@stars] and `spacetime` [@spacetimebook] by R-Spatial allows for both space and time dimension for raster and vector data, but the underlying data structure is a multi-dimensional array, which could be difficult to operate for R users who are more familiar with the operation in 2D dataframe/ tibble.

<!-- - `sp` + `spacetime` (time series data from package `xts`) -->

<!-- - `sf`  -->
<!--   - simple features geometry  -->
<!--   - only spatio -->
<!--   - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure -->

<!-- - `stars` -->
<!--   - a newer structure allow both spatio + temporal  -->
<!--   - array -->

In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [... ] workflow to work with temporal data. In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier. However, a tsibble object, by construction, always requires the `index` in its structure. This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e. features of each series) don't have the time dimension. Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects. In addition, the long form structure of a tsibble object means spatio variables (i.e. longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp. This repetition is unnecessary and would inflate the object size for long series.

# A new data structure for spatio-temporal data

Intro to cubble:

- list-column: rowwise_df with temporal variables, including the time index, nested. 
  - Focus on spatio: those output per station

- long form: grouped_df 
  - Focus on temporal

Compatible with tidyverse manipulation and tsibble

# Examples

Daily climate data (prcp, tmax, and tmin) from RNOAA - lots of stations across Australia

An exploratory data analysis questions: What's the climate profile look like in Australia

  - General features: Any general trend/ fluctuation in prcp, tmax, and tmin?
  - Local features: Any station stands out from the crowd?

## Manipulation

  - data quality check: filter out stations have variables not properly recorded 
  - data summary: 
    - daily -> monthly/ weekly, 
    - summarise by mean for tmax/ tmin, sum for prcp
  -

## Graphics

Static + interactive -> tooltip to show additional information upon hovering

  - Where are those stations on the map? 
    - Mention mostly aero, airport, and lighthouse

# Summary