### 2022-02-27 main.Rmd
chunk in changes

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolas@langrene.com
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Spatio-temporal data refers to measurements taken across space over time. In practise, spatial and temporal data many times need to be combined in order to form a spatio-temporal data set. However, these two types of data are collected at different observational units and this poses a problem on which format to use for different data wrangling tasks. Analysts can either organised the data in a single table format that combines all the variables together, or in two separate tables, each focusing on the spatial or temporal aspect of the data. Both formats are useful in exploratory data analysis, visualisation, and, modelling, however it can be labourious to constantly switch between the two formats during the analysis and write code to cast the data into the desireable shape. This paper presents a new \proglang{R} package, \pkg{cubble}, that provides a new data sturcture to ease the transformation between the two formats in spatio-temporal data. The new data structure organises spatial and temporal variables as two forms of a single data object so that they can be wrangled separately while syncronised as a whole piece. \pkg{Cubble} also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics and performing spatio-temporal transformation. The paper will also demonstrate \pkg{cubble} with examples using Australia climate weather stations, river level data, and climate reanalysis (ERA5) data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: rticles::jss_article
bibliography: references.bib
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = "figures/",
  fig.height = 10
)
options(prompt = "R> ", continue = "+ ", 
        tibble.print_max = 5, tibble.print_min = 5)
```

```{r echo = FALSE}
library(cubble)
library(forcats)
library(stringr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(GGally)
library(plotly)
library(crosstalk)
library(patchwork)
```

\newpage

```{r child = "1-intro.Rmd"}

```

```{r child = "2-the-cubble-pkg.Rmd"}

```

```{r child = "3-advanced.Rmd"}

```

```{r child = "4-1-basic.Rmd"}

```

```{r child = "4-2-agg.Rmd"}

```

```{r child = "4-3-water-level.Rmd"}

```

```{r child = "4-4-netcdf.Rmd"}

```

```{r child = "4-5-interactive.Rmd"}

```

```{r child = "5-conclusion.Rmd"}

```