### 2022-04-26 4-2-agg.Rmd

---
title: "4-2-aggre"
output: pdf_document
---

## Australia precipitation pattern in 2020

In the previous example, there has been some overlap of glyphs when stations are near each other, (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map. This example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.

```{r}
load(here::here("data/station_nested.rda"))
```

```{r echo = TRUE, eval = FALSE}
station_nested <- weatherdata::climate_full |> 
  mutate(cluster = ...)
```

To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:

```{r echo = TRUE}
cluster_nested <- station_nested |> switch_key(cluster) 
```

With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

```{r eval = FALSE, echo = TRUE}
cluster_nested <- cluster_nested |> get_centroid()
```

Long form cubbles at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data. Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
knitr::include_graphics(here::here("figures/basic-agg.png"))
```