### 2022-04-27 arxiv/cubble-arxiv.Rmd
Update abstract

---
title: "cubble: An R Package for Structuring Spatio-temporal Data"
authors:
  - name: H. Sherry Zhang
    affiliation: Monash University
    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University
    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: dicook@monash.edu
  - name: Ursula Laa
    affiliation: University of Natural Resources and Life Sciences
    location: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: BNU-HKBU United International College
    location: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolaslangrene@uic.edu.cn
  - name: Patricia Menéndez
    affiliation: Monash University
    location: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
abstract: |
    Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance instance, temporal trends of multiple variables across locations. In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling. The propose \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the  capability to handle hierarchical spatial and temporal structures. \color{green}{The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.}
keywords:
- spatial, temporal, spatio temporal, R, exploratory data analysis, environmental data, climate data
bibliography: "../references.bib"
cls: "../jss.cls"
bst: "../jss.bst"
preamble: |
  \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
link-citations: true
---
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = here::here("figures/"),
  fig.height = 10,
  root.dir = here::here()
)
options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5)
```

```{r}
library(cubble)
library(dplyr)
```

\newpage

```{r child = here::here("1-intro.Rmd")}

```

```{r child = here::here("2-0-cube.Rmd")}

```

```{r child = here::here("2-the-cubble-pkg.Rmd")}

```

```{r child = here::here("3-advanced.Rmd")}

```

```{r child = here::here("4-1-basic.Rmd")}

```

```{r child = here::here("4-2-agg.Rmd")}

```

```{r child = here::here("4-3-water-level.Rmd")}

```

```{r child = here::here("4-4-netcdf.Rmd")}

```

```{r child = here::here("4-5-interactive.Rmd")}

```

```{r child = here::here("5-conclusion.Rmd")}

```