### 2022-04-30 5-conclusion.Rmd
Updating conclusion

---
title: "5-conclusion"
output: pdf_document
---
# Conclusion {#conclude}

This paper presents an \proglang{R} package \pkg{cubble} for organising, manipulating and visualising spatio-temporal data. The package introduces a new data structure for spatio-temporal data, \code{cubble}, that connects the time invariant and varying variables and that allows the user to work with a nested and long form of the data. This design frees the data analysts from spending time on organising variables of different observational units. The data structure and the capabilities introduced in this package can be used and combined with existing spatio-temporal R packages such as \pkg{sf}, data wrangling packages such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2}, \pkg{plotly} and \pkg{leaflet}.

The data structure and functions are demonstrated with extensive examples. These include creating and coercing wild-caught data with potential mismatch on sites, handling hierarchical data, matching time series spatially and temporally, as well as reproducing ERA5 results from NetCDF data. Visualization of the \code{cubble} objects and derivatives is presented via interactive graphic pipelines using \pkg{plotly} and \pkg{leaflet}.

Future directions of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. In the spatial aspect, the simple feature representation wraps spatial coordinates in a list column, for some complex spatial operations, while sometimes, analysts may want to unpack them into longitude and latitude columns to work from. A smooth transition between the coordinate and simple feature representation can be added to \pkg{cubble}. \pkg{cubble} can also be extended to handle sites with moving coordinates.

The goal of this work is to add capabilities into the spatio-temporal practitioners toolbox that facilitates their work within the tidy data framework.

<!-- and this would involve constructing a \code{sp} column for locations, similar to \code{ts} for time, \textcolor{green}{and a form locations can be pivoted into, like the long form for temporal variables}. The multivariate aspect, \pkg{cubble} can also be extended with interface to more high dimensional visualisation methods, i.e., the tour method, to understand variable importance or comparing location similarities. -->

\newpage

# Acknowledgement

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. The article is created using \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in R. The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

# Appendix

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "An illustration of the data model under interactive graphics with cubble. When a point on the time series is selected, the corresponding row in the long cubble will be activated. This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle). Both plots will be updated with the full line selected and the point highlighted on the map (right)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

\newpage