### 2022-06-13 4-4-netcdf.Rmd
First update of seciton 5.5

---
title: "4-4-netcdf"
output: pdf_document
---

## ERA5: climate reanalysis data

The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr]. In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifially look at the information  contained on the variables *specific humidity* measuring the mass of water vapour per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. Time-wise we restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

```{r eval = FALSE, echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(raw, vars = c("q", "z"))
```

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5. It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.

```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)."}
knitr::include_graphics(here::here("figures/netcdf.png"))
```