### 2022-06-13 4-4-netcdf.Rmd
Add motivation example 5.5

---
title: "4-4-netcdf"
output: pdf_document
---

## ERA5: climate reanalysis data

The ERA5 dataset [@hersbach2020era5] provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF) and it can be directly downloaded from [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via the R package \pkg{ecmwfr} [@ecwmfr]. In this example we focus on the \code{era5-pressure} data which provides hourly data on pressure levels from 1970 to present. Here we specifically look at the information  contained on the variables *specific humidity* measuring the mass of water vapor per kilogram of moist air  and the *geopotential* variable recording the gravitational potential energy of a unit mass, at a particular location, relative to mean sea level. The goal of this example is to reproduce the break-up of the southern polar vortex that happened in late September and early October in 2022 in the stratosphere in the Southern Hemisphere. Therefore, we filter data corresponding to the Southern Hemisphere and restrict our analysis to four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Once downloaded, the data can be read into a \code{cubble} as:

```{r eval = FALSE, echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(raw, vars = c("q", "z"))
```

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5. It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to @hersbach2020era5, @simmons2020global, and @simmons2005ecmwf for more details.

```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020)."}
knitr::include_graphics(here::here("figures/netcdf.png"))
```