### 2022-06-16 4-2-agg.Rmd
use magritte pipe throughout

---
title: "4-2-aggre"
output: pdf_document
---

## Australian precipitation pattern in 2020 {#precip}

In the previous example, some overlapping of the glyphs occurred for a few nearby stations such as the pairs (151E, 34S) and (152E, 33S). This is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map. This new example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple $k$-means algorithm based on the distance matrix between stations is used to create 20 clusters. The data \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs. More advanced clustering algorithms can be used as well, as long as they provide a mapping from each station to a cluster.

```{r}
load(here::here("data/station_nested.rda"))
```

```{r eval = FALSE, echo = TRUE}
station_nested <- weatherdata::climate_full %>% mutate(cluster = ...)
```

To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:

```{r echo = TRUE}
cluster_nested <- station_nested %>% switch_key(cluster)
```

With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

```{r eval = FALSE, echo = TRUE}
cluster_nested <- cluster_nested %>% get_centroid()
```

Long form cubble at both levels can be accessed through stretching the nested form. With access to both station and cluster-level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data. Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

```{r eval = FALSE}
coords <- cbind(prcp_aus$long, prcp_aus$lat)
dist_raw <- geosphere::distm(coords, coords)

station_long <- cubble::prcp_aus %>%
  face_temporal(ts) %>%
  mutate(wk = lubridate::week(date)) %>%
  group_by(wk) %>%
  summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
  unfold(cluster)

set.seed(123)
station_nested <- station_long %>%
  face_spatial() %>%
  strip_rowwise() %>%
  mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
save(station_nested, file = here::here("data/station_nested.rda"))
```

```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
load(here::here("data/station_nested.rda"))
cluster_nested <- station_nested %>%
  switch_key(cluster) %>% 
  get_centroid()

cluster_long <- cluster_nested %>% 
  face_temporal() %>% 
  group_by(wk) %>%
  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
  unfold(cent_long, cent_lat)

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

ggplot_smooth <- cluster_long %>% 
  ggplot() +
  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)

smoother <- layer_data(ggplot_smooth) %>% 
  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))

p1 <- ggplot(data = smoother, 
             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_text(data = cluster_nested, 
            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
  geom_glyph(height = 2, width = 4) + 
  theme_void()

p2 <- ggplot() + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
                          expand = 0, radius = 0,
                          aes(x = long, y = lat, group = cluster)) + 
  theme_void()

(p1 | p2) + plot_annotation(tag_levels = 'A')
```