### 2022-06-17 jss/appendix.Rmd
appendix

---
title: "Appendix to 'cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data'"
author: "Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
date: '2022-06-14'
output: 
  pdf_document:
    number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the supplementary material for the main paper, containing additional information about the data sets used, and another illustration of linking between multiple plots.

# Additional illustration on multiple linked plots

<!-- Figure \ref{fig:illu-interactive-2} illustrates the linking pipeline from a point selection on the time series to the map. -->

This figure is a supplement to Section 4.3 of the main paper, illustrating how linking from the time series plot to the map is achieved.

```{r illu-interactive-2, echo = FALSE, fig.align="center", out.height="40%", out.width = "100%", fig.cap = "Linking between multiple plots. The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles). When a point on the time series is selected, the corresponding row in the long cubble will be activated (a). This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (b). Both plots will be updated with the full line selected and the point highlighted on the map (c)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.005.png"))
```

# Scripts for creating the example data

## Historical maximum temperature

The script below presents the codes required to obtain the data `historical_tmax` used in the example of Section 5.2 *Australian historical maximum temperature*. The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time. A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.

```{r eval = FALSE}
library(tidyverse)
library(cubble)
all_stations <- rnoaa::ghcnd_stations() %>%
  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
  filter(last_year >= 2020) %>%
  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
  select(-state, -gsn_flag) %>%
  select(id, longitude, latitude, elevation, name,
         wmo_id, element, first_year, last_year) %>%
  rename(long = longitude, lat = latitude, elev = elevation)

tmax_stations <- all_stations %>% 
  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))

raw_tmax <- all_stations %>%
  rowwise() %>%
  mutate(ts = list(rnoaa::meteo_pull_monitors(
    monitors = id, var = "TMAX",
    date_min = glue::glue("{first_year}-01-01"),
    date_max = glue::glue("{last_year}-12-31")
    ) %>%
      select(-id)
    )
  )

historical_tmax <- raw_tmax %>%
  select(-element) %>%
  unnest(ts) %>%
  mutate(tmax = tmax/10) %>%
  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
  as_cubble(index = date, key = id, coords = c(long, lat))

save(historical_tmax, file = here::here("data/historical_tmax.rda"))
```

## Australian 2016-2020 climate data

The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`. The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.

```{r eval = FALSE}
# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
aus_stations <- all_stations %>%
  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
  nest(element: last_year) %>%
  rowwise() %>%
  filter(nrow(data) == 3) %>%
  select(-data)

aus_climate_raw <- aus_stations %>%
  rowwise() %>%
  mutate(ts = list(
    rnoaa::meteo_pull_monitors(
      monitors = id, var = c("PRCP", "TMAX", "TMIN"),
      date_min = "2016-01-01", date_max = "2020-12-31"
      ) %>%
      select(-id)
    )
  )

climate_full <- aus_climate_raw %>%
  unnest(ts) %>%
  mutate(tmax = tmax/10, tmin = tmin/10) %>%
  cubble::as_cubble(key = id, index = date, coords = c(long, lat))

save(climate_full, file = here::here("data/climate_full.rda"))
```