### 2022-06-17 4-1-basic.Rmd
Move example 5.3 to appendix + update appendix

---
title: "4-examples"
output: html_document
---

# Examples {#examples}

The five examples here together with an additional in the Appendix are chosen to illustrate creating a cubble for exploring COVID data where location identities need to be matched, using spatial transformations to make a glyph map of seasonal temperature changes over years, aggregating information spatially to explore precipitation patterns, matching river level data and weather station records for analysis of water supply, reading netCDF format data to reproduce a climate reanalysis plot, and the workflow to create complex interactive linked plots.

## Victoria COVID spatio-temporal incidence and spread {#covid}

Since the start of the pandemic, the Victoria State Government in Australia has provided daily COVID counts by local government area (LGA). This data can be used to visualise COVID incidence and spread spatially, when combined with map polygon data available from the Australian Bureau of Statistics. These different sources need to be combined for the analysis, by matching the LGA names. Here is how to do this with \pkg{cubble}, including how to handle mismatches arising from different names of the same LGAs in the two tables. The COVID data is a `csv` and looks like:

```{r eval = FALSE}
library(tidyverse)
#https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
                  col_names = c("date", "postcode", "source", "lga"), skip = 1)

covid <- raw %>%
  count(date, lga, source, postcode) %>%
  filter(date >= as.Date("2022-01-01")) %>%
  filter(source == "Contact with a confirmed case") %>%
  group_by(date, lga, source) %>%
  count() %>%
  ungroup(date) %>%
  mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
  tsibble::as_tsibble(key = lga, index = date) %>%
  tsibble::fill_gaps(.full = TRUE)
save(covid, file = here::here("data/covid.rda"))
```

```{r}
load(here::here("data/covid.rda"))

lga <- strayr::read_absmap("lga2018") %>%
  rename(lga = lga_name_2018) %>%
  filter(state_name_2016 == "Victoria") 
```

```{r echo = TRUE}
covid %>% head(5)
```

and the spatial polygons are an ESRI shapefile that looks like:

```{r echo = TRUE}
lga %>% head(5)
```

The function \code{as_cubble()} is used to create a \code{cubble} object from the two spatial and temporal tables, and requires specifying the arguments (\code{key}, \code{index}, and \code{coords}) (as described in Section \@ref(create)). It will automatically try to match the sites in both tables and will show a warning message when there are mismatches, as shown below:

```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat)
  )
```

It can be seen that there are two-way mismatches -- LGAs in the COVID data that don't match with LGAs in the spatial polygon data, and vice versa. The mismatches can be identified by using the \code{output = "unmatch"} argument. <!--Note that \pkg{cubble} will attempt to pair the unmatched sites, these are shown separately from the ones that cannot be paired.-->

```{r echo = TRUE}
pair <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat),
  output = "unmatch"
  )

pair
```

With this information both tables can be fixed, to create the desired cubble, as follows:

```{r echo = TRUE}
lga <- lga %>%
  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
  filter(!lga %in% pair$others$spatial)

covid <- covid %>% filter(!lga %in% pair$others$temporal)

cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))
```

## Australian historical maximum temperature {#historicaltmax}

```{r}
load(here::here("data/historical_tmax.rda"))
```

The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The data \code{historical_tmax} is a subset containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations, extracted using the `rnoaa` package. <!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature  between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

First of all, the stations across the two states need to be subsetted from the original data  \code{historical_tmax}. For that, the country and the station identifiers, stored within the 11 digits of the \code{id} variable entries, need to be retrieved. The country code is located in the first 5 digits (Australia is represented by "ASN00") while the next 6 digits encode the station information following the [Australian Bureau of Meteorology (BOM)](http://www.bom.gov.au/climate/cdo/about/site-num.shtml) coding protocols. According to those, New South Wales stations correspond to entries in the range 46-75 and the Victorian stations are stored under 76-90. Filtering Victoria and New South Wales stations is a spatial operation and hence uses the \code{cubble} nested form:

```{r echo = TRUE}
tmax <- historical_tmax %>%
  filter(between(stringr::str_sub(id, 7, 8), 46, 90))
```

A monthly maximum average temperature is then calculated for both periods where \code{cubble} easily allows to compute this time operation across all the stations. This is a temporal operation and therefore the nested cubble needs to be switched to the long cubble form by \code{face_temporal()}:

```{r echo = TRUE}
tmax <- tmax %>%
  face_temporal() %>% 
  group_by(month = lubridate::month(date),
         group = as.factor(ifelse(lubridate::year(date) > 2015,
                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

Observing the results of the operation above show that a few stations do not have records during the time period 1971-1975. Further investigation shows that while the first and last year of each series for those stations are recorded, the years in between are not. Hence,  these stations are  located and filtered out by examining whether the summarised time series has 24 months. Here, the long cubble needs to be switched to the nested form to carry out this spatial operation using \code{face_spatial()}:

```{r echo = TRUE}
tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24)
```

Lastly, to create a glyph map where the series of each station are overlayed on a map of Australia, both the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates defining the bounding box for the plot need to be on the same table. For that, spatial variables can be moved to the long form with \code{unfold()}:

```{r echo = TRUE}  
tmax <- tmax %>% face_temporal() %>% unfold(long, lat)
```

the variable storing the maximum monthly average temperatures \code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:

```{r eval = FALSE, echo = TRUE}
nsw_vic <- ozmaps::abs_ste %>%
  filter(NAME %in% c("Victoria", "New South Wales"))

ggplot() +
  geom_sf(data = nsw_vic,
          fill = "transparent", color = "grey", linetype = "dotted") +
  geom_glyph(data = tmax,
             aes(x_major = long, x_minor = month,
                 y_major = lat, y_minor = tmax,
                 group = interaction(id, group), color = group),
             width = 1, height = 0.5) +
  ...
```

```{r create-inset, eval = FALSE}

tmax %>% filter(id == "ASN00048027") %>%
  ggplot(aes(x = month,
             y = tmax,
             color = group)) +
  geom_line(size = 1.5) +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  scale_x_continuous(breaks = seq(1, 12, 1), 
                     labels = c("J", "F", "M", "A", "M", "J", 
                                "J", "A", "S", "O", "N", "D")) +
  labs(x = "", y  = "Temp (C)", title = "ASN00048027: Cobar") +
  theme_bw() +
  theme(
    aspect.ratio = 0.3,
    axis.text = element_text(size = 20),
    title =  element_text(size = 20)
  )

#ggsave(filename = here::here("figures/basic-manip-inset.png"), width = 7)
```

```{r basic-manip, fig.height = 10, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020). The corresponding average time series for the Cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations."}
nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria", "New South Wales"))

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", color = "white") + 
  geom_glyph(data = tmax, 
             aes(x_major = long, x_minor = month, 
                 y_major = lat, y_minor = tmax,
                 group = interaction(id, group), 
                 color = group),
             width = 1, height = 0.5) +
  scale_color_brewer("", palette = "Dark2") + 
  theme_bw() + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom") +
  labs(x = "Longitude", y = "Latitude")

box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
single <- tibble::tibble(img = here::here("figures/basic-manip-inset.png"))

p1 +
  geom_rect(data = box_df,
            aes(xmin = long - 0.6, xmax = long + 0.6,
                ymin = lat - 0.12, ymax = lat + 0.35),
            fill = "transparent", color = "black") +
  ggimg::geom_point_img(
    data = single, aes(x = 143.7, y = -30, img = img), size = 6)

```
It is worth to mentioned that some overlapping of the glyphs occurred for a few nearby stations as it can be seen from Figure  \ref{fig:basic-manip}. This is a common issue when mapping multiple stations in a geographical area. This issue can be mitigated in some instances by aggregating data to group the data into a larger areas which might be useful to learn about clusters in the data as well as for visualizing those clusters in a glyph map. An continuing discussion of Example \@ref(historicaltmax) discussing all those issues and can be found in the Appendix.