### 2022-06-24 reference-hack.Rmd

---
title: "reference-hack"
output: html_document
---

The `rticles::jss_article` template used during preparation (https://github.com/rstudio/rticles/tree/c085e125d874c11bb474ec53f4b754e179f8a4e9) has the behavior that a reference prints out all the author names when it is used for the first time. See a standalone example [here](https://github.com/huizezhang-sherry/jss).

Here a hack is used to overcome this temporarily

```{r}
tex <- readLines(here::here("references.bib"))

b <- tex %>% stringr::str_extract("\\@(article|techreport|misc|Manual|Book)\\{.+,")
b <- b[!is.na(b)]
ref_head <- b %>% stringr::str_replace("\\@(article|techreport|misc|Manual|Book)\\{", "") %>% 
  stringr::str_replace(",", "")

res <- paste0("@", ref_head, collapse = ", ")
```

Here I reference `r res`