### 2022-06-27 code/appendix.Rmd
clean up repo

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(cubble)
library(patchwork)
library(leaflet)
library(leafpop)

```
```{r echo = FALSE}
load(here::here("data/station_nested.rda"))
load(here::here("data/climate_full.rda"))

```
```{r eval = FALSE}
## # The kmeans clustering in station_nested data can takes some time to run,
## # This chunk runs the algorithm and save the result as staiton_nested for further use
## coords <- cbind(prcp_aus$long, prcp_aus$lat)
## dist_raw <- geosphere::distm(coords, coords)
## 
## station_long <- cubble::prcp_aus %>%
##   face_temporal(ts) %>%
##   mutate(wk = lubridate::week(date)) %>%
##   group_by(wk) %>%
##   summarise(prcp = sum(prcp, na.rm = TRUE)) %>%
##   unfold(cluster)
## 
## set.seed(123)
## station_nested <- station_long %>%
##   face_spatial() %>%
##   strip_rowwise() %>%
##   mutate(cluster = kmeans(dist_raw,centers = 20, nstart = 500)$cluster)
## save(station_nested, file = here::here("data/station_nested.rda"))

```
```{r basic-agg, fig.height = 7, fig.width = 15, fig.cap="Profile of aggregated precipitation at 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
load(here::here("data/station_nested.rda"))
cluster_nested <- station_nested %>%
  switch_key(cluster) %>% 
  get_centroid()

cluster_long <- cluster_nested %>% 
  face_temporal() %>% 
  group_by(wk) %>%
  summarise(prcp = mean(prcp, na.rm = TRUE)) %>% 
  unfold(cent_long, cent_lat)

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

ggplot_smooth <- cluster_long %>% 
  ggplot() +
  geom_smooth(aes(x = wk, y = prcp, group = cluster), span = 0.4)

smoother <- layer_data(ggplot_smooth) %>% 
  left_join(cluster_long %>% select(cluster, cent_long, cent_lat), by = c("group" = "cluster"))

p1 <- ggplot(data = smoother, 
             aes(x_minor = x, y_minor = y, x_major = cent_long, y_major = cent_lat)) + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_text(data = cluster_nested, 
            aes(x = cent_long, y = cent_lat, label = cluster), inherit.aes = FALSE) +
  geom_glyph(height = 2, width = 4) + 
  theme_void()

p2 <- ggplot() + 
  geom_sf(data = state_map, inherit.aes = FALSE,
          color = "grey80", alpha = 0.4, linetype = 3) + 
  geom_point(data = station_nested, aes(x = long, y = lat), size = 0.5) +
  ggforce::geom_mark_hull(data = cluster_nested %>% tidyr::unnest(hull),
                          expand = 0, radius = 0,
                          aes(x = long, y = lat, group = cluster)) + 
  theme_void()

(p1 | p2) + plot_annotation(tag_levels = 'A')

```
```{r eval = FALSE}
## clean <-  climate_full %>%
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
##   face_temporal() %>%
##   mutate(month = lubridate::month(date)) %>%
##   group_by(month) %>%
##   summarise(tmax = mean(tmax, na.rm = TRUE),
##             tmin = mean(tmin, na.rm = TRUE),
##             diff = mean(tmax - tmin, na.rm = TRUE)) %>%
##   ungroup(month) %>%
##   face_spatial() %>%
##   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE)) %>%
##   face_temporal() %>%
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
##   unfold(name)
## 
## df_id <- clean$id %>% unique()
## p <- purrr::map(1:length(df_id), function(i){
##   dt <- clean %>% filter(id == df_id[i])
##   dt %>%
##     ggplot(aes(x = dummy_date, ymin = tmin, ymax = tmax, group = id)) +
##     geom_ribbon(color = "grey50", fill = "grey50", alpha = 0.5) +
##     theme_bw() +
##     scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##     labs(x = "Month", y = "Temperature") +
##     ylim(-10, 50) +
##     ggtitle(paste0(stringr::str_to_title(unique(dt$name))))
## })
## 
## nested <- face_spatial(clean)
## domain <- nested$temp_diff_var
## pal <- colorNumeric(
##   colorspace::sequential_hcl(
##     "Rocket",  n = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30),
##   domain = domain)
## 
## leaflet(nested) %>%
##   addTiles() %>%
##   addCircleMarkers(color = ~pal(domain), radius = 0.1,
##                    fillOpacity = 1, opacity = 1,
##                    popup = popupGraph(p, width = 300, height =200))

```
```{r interactive-popup, fig.align='center', echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 except the temperature variation is now shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup.png"))

```
```{r eval = FALSE}
## library(tidyverse)
## library(cubble)
## all_stations <- rnoaa::ghcnd_stations() %>%
##   filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
##   filter(last_year >= 2020) %>%
##   mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
##   select(-state, -gsn_flag) %>%
##   select(id, longitude, latitude, elevation, name,
##          wmo_id, element, first_year, last_year) %>%
##   rename(long = longitude, lat = latitude, elev = elevation)
## 
## tmax_stations <- all_stations %>%
##   filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))
## 
## raw_tmax <- all_stations %>%
##   rowwise() %>%
##   mutate(ts = list(rnoaa::meteo_pull_monitors(
##     monitors = id, var = "TMAX",
##     date_min = glue::glue("{first_year}-01-01"),
##     date_max = glue::glue("{last_year}-12-31")
##     ) %>%
##       select(-id)
##     )
##   )
## 
## historical_tmax <- raw_tmax %>%
##   select(-element) %>%
##   unnest(ts) %>%
##   mutate(tmax = tmax/10) %>%
##   filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>%
##   as_cubble(index = date, key = id, coords = c(long, lat))
## 
## save(historical_tmax, file = here::here("data/historical_tmax.rda"))

```
```{r eval = FALSE}
## # all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
## aus_stations <- all_stations %>%
##   filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
##   nest(element: last_year) %>%
##   rowwise() %>%
##   filter(nrow(data) == 3) %>%
##   select(-data)
## 
## aus_climate_raw <- aus_stations %>%
##   rowwise() %>%
##   mutate(ts = list(
##     rnoaa::meteo_pull_monitors(
##       monitors = id, var = c("PRCP", "TMAX", "TMIN"),
##       date_min = "2016-01-01", date_max = "2020-12-31"
##       ) %>%
##       select(-id)
##     )
##   )
## 
## climate_full <- aus_climate_raw %>%
##   unnest(ts) %>%
##   mutate(tmax = tmax/10, tmin = tmin/10) %>%
##   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
## 
## save(climate_full, file = here::here("data/climate_full.rda"))
```