### 2022-06-27 code/main.Rmd

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.path = here::here("figures/")
)

options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5,
        width = 70)

```
```{r echo = FALSE}
library(cubble)
library(tidyverse)
library(patchwork)

```
```{r }
cubble::climate_flat

```
```{r echo = TRUE}
cubble_nested <- climate_flat |>
  as_cubble(key = id, index = date, coords = c(long, lat))
cubble_nested

```
```{r echo = TRUE}
cubble_nested |>
  mutate(rain_day = sum(ts$prcp != 0))

```
```{r echo = TRUE}
cubble_long <- cubble_nested |> face_temporal()
cubble_long

```
```{r echo = TRUE}
cubble_back <- cubble_long |> face_spatial()
cubble_back
identical(cubble_nested, cubble_back)

```
```{r echo = TRUE}
cubble_unfold <- cubble_long |> unfold(long, lat)
cubble_unfold

```
```{r echo = TRUE}
climate_flat_ts <- climate_flat |> 
  tsibble::as_tsibble(key = id, index = date)
climate_flat_cb <-  climate_flat_ts |> 
  cubble::as_cubble(coords = c(long, lat))
climate_flat_cb

```
```{r echo = TRUE}
# add station-based features in the nested form.
climate_flat_cb |> 
  mutate(
    fabletools::features(
      ts, tmax, list(tmax_mean = mean, tmax_var = var)
      )
  )

```
```{r eval = FALSE}
## library(tidyverse)
## #https://www.coronavirus.vic.gov.au/victorian-coronavirus-covid-19-data
## raw <- read_csv(here::here("data/NCOV_COVID_Cases_by_LGA_Source_20220324.csv"),
##                   col_names = c("date", "postcode", "source", "lga"), skip = 1)
## 
## covid <- raw %>%
##   count(date, lga, source, postcode) %>%
##   filter(date >= as.Date("2022-01-01")) %>%
##   filter(source == "Contact with a confirmed case") %>%
##   group_by(date, lga, source) %>%
##   count() %>%
##   ungroup(date) %>%
##   mutate(roll_mean = zoo::rollmean(n, k = 14, fill = NA)) %>%
##   tsibble::as_tsibble(key = lga, index = date) %>%
##   tsibble::fill_gaps(.full = TRUE)
## save(covid, file = here::here("data/covid.rda"))

```
```{r }
load(here::here("data/covid.rda"))

lga <- strayr::read_absmap("lga2018") %>%
  rename(lga = lga_name_2018) %>%
  filter(state_name_2016 == "Victoria")

```
```{r echo = TRUE}
covid %>% head(5)

```
```{r echo = TRUE}
lga %>% head(5)

```
```{r echo = TRUE, warning = TRUE, message= TRUE}
cb <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat)
  )

```
```{r echo = TRUE}
pair <- as_cubble(
  list(spatial = lga, temporal = covid),
  key = lga, index = date, coords = c(cent_long, cent_lat),
  output = "unmatch"
  )

pair

```
```{r echo = TRUE}
lga <- lga %>%
  mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
         lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) %>%
  filter(!lga %in% pair$others$spatial)

covid <- covid %>% filter(!lga %in% pair$others$temporal)

cb <- as_cubble(data = list(spatial = lga, temporal = covid),
                key = lga, index = date, coords = c(cent_long, cent_lat))

```
```{r }
load(here::here("data/historical_tmax.rda"))

```
```{r echo = TRUE}
tmax <- historical_tmax %>%
  filter(between(stringr::str_sub(id, 7, 8), 46, 90))

```
```{r echo = TRUE}
tmax <- tmax %>%
  face_temporal() %>% 
  group_by(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))

```
```{r echo = TRUE}
tmax %>% 
  face_spatial() %>% 
  mutate(n = nrow(ts)) %>%
  arrange(n) %>%
  pull(n) %>% 
  head(10)

tmax <- tmax %>% 
  face_spatial() %>% 
  filter(nrow(ts) == 24) %>%
  face_temporal()

```
```{r }
cobar <- tmax %>% filter(id == "ASN00048027") %>%
  ggplot(aes(x = month, y = tmax, color = group)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  scale_x_continuous(
    breaks = seq(1, 12, 1), 
    labels = c("J", "F", "M", "A",
               "M", "J", "J", "A",
               "S", "O", "N", "D")
    ) +
  labs(x = "", y  = "Temp (C)") +
  theme_bw() + 
  theme(aspect.ratio = 0.3)

```
```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales for the periods (1971-1975, 2016-2020). The corresponding average time series for the Cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in New South Wales."}
tmax <- tmax %>% unfold(long, lat)
box_df <- tmax %>% face_spatial() %>% filter(id == "ASN00048027")
nsw_vic <- ozmaps::abs_ste %>% filter(NAME %in% c("Victoria","New South Wales"))

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white", size=2) + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

(p1 | cobar)  + 
  patchwork::plot_layout(width = c(1,1)) + 
  plot_annotation(tag_levels = 'a')

```
```{r }
load(here::here("data/climate_full.rda"))

climate <- climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  face_temporal() %>%
  filter(lubridate::year(date) == 2020) %>%
  face_spatial() %>%
  mutate(type = "climate")

river <- river %>% mutate(type = "river")

```
```{r echo = TRUE}
res <- match_sites(
  river, climate,
  temporal_by = c("Water_course_level" = "prcp"),
  temporal_independent = "prcp",  
  temporal_n_highest = 30,
  temporal_min_match = 15, 
)

```
```{r }
res %>% print()

```
```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
vic_map <- ozmaps::abs_ste %>% 
  filter(NAME == "Victoria") %>% 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate), 
             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
  geom_point(data = res, aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res %>% filter(type == "river"), aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

res_long <- res %>%
  face_temporal(ts) %>%
  unfold(group, type) %>%
  rename(prcp = matched_var) %>% 
  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))

p2 <- res_long %>% 
  ggplot(aes(x = date, y = prcp, color = type, group = id)) +
  geom_line(alpha = 0.8) +
  facet_wrap(vars(group)) +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  theme_bw() +
  scale_x_date(date_labels = "%b") + 
  labs(x = "Date", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom")

```
```{r echo = TRUE}
raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
dt <- as_cubble(
  raw, vars = c("q", "z"),
  long_range = seq(-180, 180, 1), lat_range = seq(-88, 88, 1))

```
```{r eval = FALSE}
## # prepare data/era5-pressure.nc
## library(ecmwfr)
## my_request <- list(
##   product_type = "reanalysis",
##   format = "netcdf",
##   pressure_level = "10",
##   year = "2002",
##   month = c("09", "10"),
##   day = c("04", "22", "26", "30"),
##   time = "00:00",
##   area = c(-5, -180, -90, 180),
##   variable = c("specific_humidity", 'geopotential'),
##   dataset_short_name = "reanalysis-era5-pressure-levels",
##   target = "era5-pressure.nc"
## )
## #
## # # here you need to create an account on Climate Data Store
## # # at https://cds.climate.copernicus.eu/cdsapp#!/home) to obtain the user ID and API key
## # # then set up the with ecmwfr::wf_set_key(user, key, service = "cds")
## # # the user below is the same `user` in the setup line
## wf_request(
##   user = "113339",
##   request = my_request,
##   transfer = TRUE,
##   path = here::here("data/"))

```
```{r netcdf, out.width="100%", fig.height = 4, fig.width = 10, fig.cap = "A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high speicfic humidity, splits into two on 2002-09-26 and furthers split into four on 2002-10-04."}
date <- c("2002-09-22", "2002-09-26", "2002-09-30", "2002-10-04") %>% as.Date()
res <- dt %>% 
  face_temporal() %>% 
  filter(lubridate::date(time) %in% date) %>%
  unfold(long, lat) %>% 
  mutate(q =  q* 10^6)

box = c(xmin = -180, ymin = -90, xmax = 180, ymax = -5)
sf::sf_use_s2(FALSE)
country <- rnaturalearth::ne_coastline("small", returnclass = "sf") %>% 
  sf::st_crop(box) %>% 
  sf::st_cast("MULTILINESTRING")

res %>% 
  ggplot() +
  # q for specific humidity
  geom_point(aes(x = long, y = lat, color = q), size = 2) + 
  #geom_tile(aes(x = long, y = lat, fill = q), color = "transparent") +
  # z for geopotential
  geom_contour(data = res, aes(x = long, y = lat, z = z),
               color = "grey20", binwidth = 4000, size = 0.5) +
  geom_sf(data = country , alpha = 0.5, fill = "transparent", color = "lightgreen") +
  coord_sf(
    crs = "+proj=stere +lat_0=-90 +lon_0=-180 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ", default_crs = sf::st_crs(4326), label_graticule = "EW") + 
  theme_bw() + 
  theme(legend.position = "bottom",
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        axis.text = element_blank(), 
        panel.grid = element_blank()) + 
  facet_wrap(vars(as.Date(time)), nrow = 1, shrink = TRUE)+  
  colorspace::scale_color_continuous_sequential("Purple-Yellow", name = "Specific humidity") + 
  theme(legend.text = element_text(size = 13),
        legend.title = element_text(size = 13)) + 
  labs(x ="", y = "")

```
```{r eval = FALSE}
## # This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
## library(crosstalk)
## library(plotly)
## 
## clean <- climate_full %>%
##   filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040",
##                     "ASN00097053", "ASN00040856", "ASN00015548")) %>%
##   face_temporal() %>%
##   mutate(month = lubridate::month(date)) %>%
##   group_by(month) %>%
##   summarise(
##     tmax = mean(tmax, na.rm = TRUE),
##     tmin = mean(tmin, na.rm = TRUE),
##     diff = mean(tmax - tmin, na.rm = TRUE)
##     ) %>%
##   ungroup(month) %>%
##   mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) %>%
##   face_spatial() %>%
##   mutate(temp_diff_var = ts %>% pull(diff) %>% var(na.rm = TRUE))
## 
## nested <- clean %>%
##   SharedData$new(~id, group = "cubble")
## 
## long <- clean %>%
##   face_temporal() %>%
##   unfold(temp_diff_var) %>%
##   arrange(temp_diff_var) %>%
##   SharedData$new(~id, group = "cubble")
## 
## state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
## 
## p1 <- nested %>%
##   ggplot() +
##   geom_sf(data = state_map, fill = "grey95", color = "white") +
##   geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
##   colorspace::scale_color_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   labs(x = "Longitude", y = "Latitude") +
##   coord_sf() +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "none"
##     )
## 
## p2 <- long %>%
##   ggplot(aes(x = dummy_date, group = id,
##              fill = temp_diff_var, color = temp_diff_var)) +
##   geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) +
##   geom_line(aes(y = tmin), alpha = 0.1) +
##   geom_line(aes(y = tmax), alpha = 0.1) +
##   geom_point(aes(y = tmax), size = 0.1) +
##   geom_point(aes(y = tmin), size = 0.1) +
##   colorspace::scale_fill_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   colorspace::scale_colour_continuous_sequential(
##     "Rocket",  n_interp = 7, cmax = 90, rev = TRUE,
##     c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
##   scale_x_date(date_labels = "%b", date_breaks = "1 month") +
##   labs(x = "Month", y = "Temperature") +
##   theme_bw() +
##   theme(
##     panel.grid.major = element_blank(),
##     legend.position = "bottom"
##     )
## 
## out <- bscols(
##   ggplotly(p1, width = 900, height = 700, tooltip = "label") %>%
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.1),
##   ggplotly(p2, width = 1200, height = 700, tooltip = "label") %>%
##     highlight(
##       on = "plotly_selected",
##       off = "plotly_deselect",
##       opacityDim = 0.012),
##   widths = c(5, 5)
## )
## 
## htmltools::save_html(out, file = "figures/linking.html")
##

```
```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row. The maximum and minimum temperatures are shown using a ribbon. The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo AWS). Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))

```
```{r}
sessioninfo::session_info()
```