### 2022-06-27 4-other-features.Rmd

---
title: "3-advanced"
output: pdf_document
---

# Other features and considerations {#others}

## Hierarchical structure

Spatial locations can have grouping structures either inherent to the data (e.g. state within country) or obtained during the analysis (e.g. cluster id). In this case, it can be useful to summarize variables at various levels of the hierarchy. The function \code{switch_key()} can be used to change the grouping level of spatial locations. The diagram in Figure \ref{fig:illu-hier} shows how this function can be used to switch the grouping from station ids to cluster ids. The result can also be stretched into long form. By specifying

\code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)},

the cubble redefines the cubble key from the \code{id} column in \code{station_nested} to the \code{cluster} column in \code{cluster_nested}. All the spatial variables belonging to the \code{cluster} column are now nested into a \code{.val} column which allows for summarizing based on cluster.

```{r illu-hier, echo = FALSE, fig.align="center",out.height="40%", out.width = "100%", fig.cap = "Hierarchical spatial structure can be handled using \\code{switch\\_key()}, to create summaries based on any level. Here the switch is between the station id and a cluster id. Once the change is made the data can be stretched into the long form."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.003.png"))
```

## Data fusion and matching {#matching}

One task that may interest spatio-temporal analysts is combining data collected at nearby but not exactly the same sites, for example, weather station measured rainfall and river levels. This can be considered to be a matching problem [@stuart2010matching; @mcintosh2018using] to pair similar time series from nearby locations, or even a data fusion exercise that merges data collected from different sources [@cocchi2019data]. The function \code{match_sites()} in \pkg{cubble} provides a simple algorithm for this task. The algorithm first matches spatially by computing the pairwise distance on latitude and longitude. Then it matches temporally by computing the number of matched peaks within a fixed length moving window. Figure \ref{fig:illu-matching} illustrates this temporal matching. In the two series, \code{A} and \code{a}, three peaks have been identified in each. An interval, of fixed length, is constructed for each peak in series \code{A}, while the peaks in series \code{a} are tested against whether they fall into any of the intervals. Here two out of three peaks match. Options for \code{match_sites()} are:

```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series \\code{A}. Two peaks in series \\code{a} fall into the intervals and hence the two series are considered to have two matches."}
knitr::include_graphics(here::here("figures/illu-matching.png"))
```

 - \code{spatial_n_keep}: the number of spatial match for each site to keep
 - \code{spatial_dist_max}: the maximum distance allowed for a matched pair
 - \code{temporal_n_highest}: the number of peaks used - 3 in the example above
 - \code{temporal_window}: the length of the interval - 5 in the example above
 - \code{temporal_min_match}: the minimum number of matched peaks for a valid matched pair

## Interactive graphics

<!-- Interactive graphics can listen to users' actions on the plot to provide additional information that facilitates data exploration. This is a useful technique for spatio-temporal data since users can zoom or pan the map to view the local and global structure of the map; use tooltips or popups to query more information about a graphic element; or highlight points to explore its linked views in other plots. In the R community, many implementations have been developed to connect \proglang{R} to \proglang{javascript} to create interactive graphics. In relation to spatio, temporal, and spatio-temporal data, the general purpose packages \pkg{plotly} [@plotly] and \pkg{leaflet} [@leaflet] realise various interactive actions through their corresponding javascript libraries. \pkg{crosstalk} [@crosstalk] and \pkg{tsibbletalk} [@tsibbletalk] implement brushed linking between htmlwidgets. \pkg{ggiraph} [@ggiraph] enables tooltip, self-linking, and customised actions specified through its own \proglang{javascript}. -->

<!-- While many graphic implementations present worked examples to illustrate the usage of the package, few documents the underlying pipeline that transforms the raw data step-by-step to the final view on the screen. There have been some early work in building the data pipeline for (interactive) graphics [@buja1988elements; @buja1996interactive;  @sutherland2000orca] and more recent discussions include @wickham2009plumbing, @xie2014reactive, and @cheng2016enabling. -->

The cubble workflow fits works well with an interactive graphics pipeline (e.g. @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in R with the package crosstalk [@crosstalk]. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series using a cubble. The map (produced from the nested form) and time series (produced from the long form) are both shared crosstalk objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested cubble object, which is then communicated to the long cubble -- all the observations with the same \code{id} (b) will be selected. The long cubble will then highlight the corresponding series in the time series plot (c).

Linking is also available starting from the time series plot, by selecting points. This will be activate rows having the same \code{id} in the long cubble. The corresponding rows in the nested cubble are activated, and highlighted the map. (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable. <!--\ref{fig:illu-interactive-2} in the Appendix illustrates this process.-->

```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots. The line plots and the map are constructed from shared crosstalk objects (long and nested cubbles). When a station is selected on the map (a), the corresponding row in the nested cubble will be activated. This will link to all the rows with the same id in the long cubble (b) and update the line plot (c)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.004.png"))
```

## Spatio-temporal transformations {#st_transformation}

Spatio-temporal data lends itself to a range of transformations. Glyph maps (Section \ref{unfold}) transform the measured variable and time coordinates into microplots at the spatial locations. Calendar plots [@wang2020calendar] deconstruct time to produce plots of variables in a calendar format. Summarizing multiple variables is commonly done using projections, or linear combinations. Here we elaborate on the transformations made to produce a glyph map.

The \proglang{R} package \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equations 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

```{r eval = FALSE, echo = TRUE}
library(ggplot2)
gly <- glyphs(data,
              x_major = ..., x_minor = ...,
              y_major = ..., y_minor = ..., ...)

ggplot(gly, aes(gx, gy, group = gid)) +
  geom_path()
```

A new implementation of the glyph map as a ggproto, \code{GeomGlyph}, has been made in the \pkg{cubble} package so that a glyph map can be created with \code{geom_glyph()}:

```{r eval = FALSE, echo = TRUE}
ggplot(data = data) +
  geom_glyph(aes(x_major = ..., x_minor = ...,
                 y_major = ..., y_minor = ...))
```

An example using a glyph map is shown in Section \@ref(historicaltmax).

Some useful controls over the glyph map are also available in the \code{geom_glyph()} implementation. Polar coordinate glyph maps are specified using \code{polar = TRUE}, and parameters \code{width} and \code{height} can be specified in either absolute or relative value. Global and local scale is specified with \code{global_rescale}, which defaults to \code{TRUE}. Reference boxes and lines can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()} lines.