### 2023-05-22 3-the-cubble-pkg.Rmd
combine section 2 and 3; rearrange section 3 with bullet points

---
title: "2-the-cubble-pkg"
output: pdf_document
---

# The cubble package {#cubble}

## The cubble object

*the conceptual framework + nested and long form as per the design vignette*

  - diagram, 
  - two forms: long/ nested - spatial/ temporal
  - attributes
  - headers

<!-- Spatio-temporal data can be conceptualized using a cubical data model with three axes which typically are, time, latitude and longitude. This abstraction can be useful for generalizing operations and visualization purposes: @lu_multidimensional_2018 shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualization based on space-time cube operations. Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time". Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualizations, still applies to our spatio-temporal data manipulation and visualization approaches. -->

<!-- The most common space-time cube uses the three axes, time, latitude, longitude, and can be considered stacking space across time. Ours is a multivariate spatio-temporal cube with the three axes defined to be time, site and variables, as illustrated in the leftmost column of Figure \ref{fig:cube-def}. The time axis is the same in both versions, while the site axis now captures both latitude and longitude. Finally, variables are stacked on this space-time canvas, with one observation per site and time point. This notion is adopted to avoid using hyper-cubes when describing multivariate spatio-temporal data and is the conceptual framework behind the \code{cubble} class. With this conceptual model, operations on spatio-temporal data can be mapped to operations on the cube and the rest of Figure \ref{fig:cube-def} show examples of slicing on site, time, and variable.  -->

The \code{cubble} class is an S3 class [@advr] built on the \code{tibble} class, specifically to organize spatio-temporal data. The \code{cubble} class uses an attribute "form",  to arrange the spatial or temporal data components tidily. The form attribute can take a value of either "nested" or "long". The nested \code{cubble} is a subclass of rowwise \code{tibble} (\code{rowwise_df}). It arranges each spatial site in a row, and uses list columns to store the feature geometry and the temporal information. The long \code{cubble} is a subclass of grouped \code{tibble} (\code{grouped_df}), which expands the temporal information into the long form and stores the spatial information in a "spatial" attribute. The main functions in the package are \code{as_cubble()}, \code{face_spatial()}, \code{face_temporal()}, and \code{unfold()}.

The data \code{climate_flat} is used to illustrate functionality in the \pkg{cubble} package. This is a subset from National Oceanic and Atmospheric Administration (NOAA) [@noaa] Global Historical Climatology Network (GHCN) Daily data. It contains spatial variables, station id, latitude, longitude, elevation, station name, world meteorology organisation id, in additional to daily temporal information,  maximum and minimum temperature values and precipitation records for year 2020. The first five rows of the data are shown below:

```{r}
cubble::climate_flat
```

## Casting into the cubble class {#create}

  - key, index, coords
  - import different data: an expand data, spatial + temporal, netcdf, stars, sftime

The function \code{as_cubble()} is used to create a \code{cubble} object with three arguments: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order (longitude, latitude). The arguments \code{key} and \code{index} follow the conventions in the \pkg{tsibble} package to describe the temporal order and multiple series while \code{coords} specifies the spatial location of each site. The code below creates a \code{cubble} object out of \code{climate_flat} (a single \code{tibble}) with \code{id} as the key, \code{date} as the index, and \code{c(long, lat)} as the coordinates:

```{r echo = TRUE}
cubble_nested <- climate_flat |>
  as_cubble(key = id, index = date, coords = c(long, lat))
cubble_nested
```

Printing a \code{cubble} object provides some information about the data. Here \code{id} is the variable name to identify each location and there are five unique locations. The bounding box is \code{[115.97, -32.94, 133.55, -12.42]} and provides information about the coordinates in the data. The third row shows the name and type of all variables nested in the \code{ts} column. In this example, it includes \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

The created \code{cubble} is a subclass of the \code{rowwise_df} class where each row forms a group. All the temporal variables are nested in a list column, hence it is also called the nested \code{cubble}. The rowwise structure makes it simpler to operate on the list using the \code{mutate()} syntax, which is simpler than the \code{purr::map()} when working with a list column. For example, calculating the number of rainy days can be done by:

```{r echo = TRUE}
cubble_nested |>
  mutate(rain_day = sum(ts$prcp != 0))
```

A \code{cubble} class can be created from various common spatio-temporal data formats, including \code{tibble}, \code{tsibble}, and \code{sf} when both spatial and temporal information are available. Section \@ref(covid) describes converting multiple tables into a \code{cubble} object and Section \@ref(netcdf) illustrates how to convert a NetCDF object.

*netcdf*

The NetCDF data is another format commonly used for storing spatio-temporal data. It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid. Attributes  can be associated with dimensions or variables. Because there can be many different styles of representing this information there is a metadata convention [@hassell_data_2017] to standardize the format of the attributes. A few packages in \proglang{R} exist for manipulating NetCDF data and these include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls a C-interface: \pkg{RNetCDF} [@rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

The \pkg{cubble} package provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble} object. It maps each combination of longitude and latitude into an \code{id} as the \code{key}:

```{r echo = TRUE, eval = FALSE}
dt <- as_cubble(my_ncdf, vars = c("q", "z"))
```

Sometimes NetCDF data can be quite large, and it is best to subset the data when converting to a \code{cubble} object. We would recommend reducing to about 300 $\times$ 300 grid points for three daily variables in one year. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution.
The size of spatial grid can be reduced if longer time periods or more variables are needed, through the arguments \code{long_range} and \code{lat_range}:

```{r eval = FALSE, echo = TRUE}
dt <- as_cubble(
  my_ncdf, vars = c("q", "z"),
  long_range = seq(-180, 180, 1), lat_range = seq(-90, 90, 1)
  )
```

## Functions and methods in cubble

*Print all the methods with methods(class = "cubble_df")*

### cubble specific methods

The three cubble specific functions are \code{face_temporal()}, \code{face_spatial()}, and \code{unfold}.

The nested form can be used for those operations where the output is only indexed by the spatial identifier (\code{key}), but becomes inadequate when outputs need both a spatial and a temporal identifier (\code{key} and \code{index}). The \code{cubble} class also provides a long form, which expands the \code{ts} column and stores the spatial variables as an attribute. The function \code{face_temporal()} is used to switch from the nested \code{cubble} into the long one. The first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face. This code switches the \code{cubble} object just created into its long form:

```{r echo = TRUE}
cubble_long <- cubble_nested |> face_temporal()
cubble_long
```

The first line in the header now shows it in the long form and the third line has been changed to display the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. Unlike the nested form, the long \code{cubble} is built from a \code{grouped_df} class where all the observations from the same site form a group.

```{r face, echo = FALSE, fig.align="center", out.width = "100%", fig.cap = "An illustration of function \\code{face\\_temporal} and \\code{face\\_spatial}. In the first row, \\code{face\\_temporal} switches a \\code{cubble} object from the nested form into the long form and the focus has switched from the spatial aspect (the side face) to the temporal aspect (the front face). In the second row, \\code{face\\_spatial} switches a \\code{cubble} object back to the nested form from the long form and shifts focus back to the spatial aspect."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

Wrangling spatio-temporal data can be seen as an iterative process in the spatial and temporal dimensions. Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}. The second row of Figure \ref{fig:face} illustrates the function, which is used as follows:

```{r echo = TRUE}
cubble_back <- cubble_long |> face_spatial()
cubble_back
identical(cubble_nested, cubble_back)
```

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps [@Wickham2012-yr]. (How to make glyph maps will be explained in Section \ref{st_transformation}, and are illustrated in the second example.) This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame. Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long \code{cubble}:

```{r echo = TRUE}
cubble_unfold <- cubble_long |> unfold(long, lat)
cubble_unfold
```

### dplyr methods

The \pkg{dplyr} package has many tools for wrangling tidy data,  many of which are useful in the spatio-temporal analysis. The \pkg{cubble} package provides methods that support the use of the following operations in the \pkg{dplyr} package on both the nested and long forms: `mutate`, `filter`, `summarise`, `select`, `arrange`, `rename`, `left_join`, and the slice family (`slice_*`).

## Compatibility with other packages

\subsubsection[]{\code{tsibble}}

The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient. A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package. A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument `coords`:

```{r echo = TRUE}
climate_flat_ts <- climate_flat |> 
  tsibble::as_tsibble(key = id, index = date)
climate_flat_cb <-  climate_flat_ts |> 
  cubble::as_cubble(coords = c(long, lat))
climate_flat_cb
```

When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on these elements. For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:

```{r echo = TRUE}
climate_flat_cb |> 
  mutate(fabletools::features(
    ts, tmax, list(tmax_mean = mean, tmax_var = var)
    ))
```

\subsubsection[]{\code{sf} (\code{s2})}

*add an example on changing CRS?*

The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). The package \pkg{sf} provides functions that operate efficiently on this spatial information. A \code{cubble} object can store spatial information in  the \code{sf} class. Methods for the \code{sfc} class can be applied in the nested form of the cubble object.  An illustration is in Section \@ref(covid). The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.

<!-- https://www.jstatsoft.org/style#my-latex-paper-does-not-compile-when-there-is-jss-markup-in-section-titles-what-should-i-do -->

## Comparison to other spatio-temporal classes {#tidyverse}

Some readers may question why a new data structure is needed rather than directly creating a list-column  on the combined data using \code{dplyr::nest_by()}. The reason is that the \code{cubble} object is specifically designed to utilize the spatio-temporal structure when arranging observations in a single object. Moreover, it enables easy pivoting between purely spatial, purely temporal, or unfolded into a combined form.

From the initial \pkg{spacetime} [@spacetime] package. New packages, such as \pkg{stars} [@stars] and \pkg{sftime} [@sftime], have [...]. This section compares and contrast the cubble data structure with other existing alternative, namely \pkg{stars} and \pkg{sftime}.

The `stars` package creates a `stars` object from an array and a dimension object. In many cases, i.e. satellite images, this is convenient, however, there are also many spatio-temporal data are formatted in 2D data frames.

create a dimension object, populate the data into the array

- array in \proglang{R} allows for character, 
- specific datetime class

*make the example simpler*

  - `temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5)`

This happens when different stations have different start date and require other software to reshape the data correctly.

<!-- The temporal data contains two variables are recorded at two stations (`id`) from 2023 March 21st to 2023 March 25th. Notice is that id2 only records from 2023 March 22nd and variable 2 has an `NA` value on  2023 March 23rd at id2. -->

<!-- ```{r echo = FALSE} -->
<!-- library(stars) -->
<!-- spatial <- tibble( -->
<!--   id = c(1,2), -->
<!--   long = c(130, 140), -->
<!--   lat = c(-38, -44), -->
<!--   name = c("station 1", "station 2"))  -->

<!-- spatial -->

<!-- temporal <- tibble( -->
<!--   id = c(rep(1, each = 5), rep(2, each = 4)), -->
<!--   time = c(seq(as.Date("2023-03-21"), as.Date("2023-03-25"), 1), -->
<!--            seq(as.Date("2023-03-22"), as.Date("2023-03-25"), 1)), -->
<!--   var1 = c(3, 6, 8, 2, 5, 7, 9, 14, 5), -->
<!--   var2 = c(1, 3, 4, 2, 6, 7, NA, 5, 8) -->
<!-- ) -->
<!-- temporal -->
<!-- ``` -->

<!-- The following code can be used to create the dimension object: -->

<!-- ```{r echo = TRUE} -->
<!-- spatial_sf <- spatial %>% sf::st_as_sf(coords = c("long", "lat"), crs = 4326)  -->
<!-- station_vec <- st_geometry(spatial_sf) -->
<!-- time_vec <- unique(temporal$time) -->
<!-- var_vec <- spatial_sf$name -->
<!-- d <- st_dimensions(station = station_vec, time = time_vec) -->
<!-- d -->
<!-- ``` -->

<!-- However, the issue comes when casting the variables into an array of the same dimension: -->

<!-- ```{r echo = TRUE, eval = FALSE} -->
<!-- temporal$var1 %>% matrix(byrow = TRUE, nrow = 2, ncol = 5) %>% array(dim = dim(d)) -->
<!-- ``` -->

<!-- Due to the missingness on 2023 March 21st for id2, the array reuses the first value of the variable to fill in the dimension (the missingness at (2,1) is not flagged and 3 is reused to fill in the dimension). -->

<!-- This causes an issue when multiple stations with different recording periods present in the data, commonly found in the weather data.  -->

The approach taken by the \pkg{sftime} package is to combine the spatial and temporal variables together into a joint table. [for sf objects with daily observation, memory efficiency]

Consider the climate data used in the \pkg{cubble} package, \code{climate_aus}. It contains daily precipitation, minimum and maximum temperature for 639 weather stations across Australia in 2020. The converted sftime object is `r library(cubble); object.size(sftime::st_as_sftime(climate_aus))/object.size(climate_aus)` times the size of the cubble object:

```{r echo = TRUE}
climate_aus %>% sftime::st_as_sftime() %>% object.size()
climate_aus %>% object.size()
```