### 2023-06-10 4-3-matching.Rmd
matching code update and misc

---
title: "4-3-water-level"
output: html_document
---

## River levels and rainfall in Victoria

River level and rainfall data for the same areas should have some similarity. Here we examine the river gauge data (\code{Water_course_level}) from the Bureau of Meteorology [@bom] in relation to weather station rainfall from NOAA's climate data (\code{climate}). The goal is to match water gauges with nearby weather stations, spatially and temporally, using the \code{match_sites()} function.
<!--The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges. The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature. We expect that the water level should be related to the precipitation available from climate data, since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} shows the location of available weather stations and water gauges in Victoria, and we will try to match water gauges with nearby weather stations, taking into account both spatial and temporal information from 2020.-->

```{r}
climate_vic <- climate_aus |>
  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
  mutate(type = "climate")

river <- cubble::river %>% mutate(type = "river")

vic_map <- ozmaps::abs_ste %>%  filter(NAME == "Victoria")

# ggplot() + 
#   geom_sf(data = vic_map) + 
#   geom_point(data = dplyr::bind_rows(river, climate_vic), 
#              aes(x = long, y = lat, color = type)) + 
#   theme_bw() +
#   theme(legend.position = "bottom") +
#   labs(x = "Longitude", y = "Latitude") + 
#   scale_color_brewer(palette = "Dark2")
```

This function requires passing the major and minor data sets used for matching, in this case those are \code{river} and \code{climate}.  The variables used for the temporal matching  are \code{Water_course_level} from the \code{river} data set and \code{prcp} in the climate data set. The rest of the arguments, as explained in Section \@ref(matching), correspond to the maximum and minimum number of peaks in the time series to be matched. In this example those are set to be a maximum of 30 and a minimum of 15 (approximately 2 and 1 per month).

```{r echo = TRUE,  tibble.print_max = 10,   tibble.print_min = 10}
(res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10))
```

```{r echo = FALSE}
res_sp <- res_sp %>% 
  filter(!group %in% c(5, 8)) %>% 
  tidyr::pivot_longer(from:to, names_to = "order", values_to = "id") %>% 
  select(-order)
res_sp <- bind_rows(climate_vic, river %>% rename(id = station)) %>% 
  inner_join(res_sp) 
```

```{r}
res_sp
```

```{r echo = TRUE}
(res_tm <- res_sp %>% 
  match_temporal(
    match_id = group,
    temporal_by = c("prcp" = "Water_course_level"),
    temporal_n_highest = 30,
    temporal_min_match = 15))
```

Or use \code{return_cubble = TRUE} to return a cubble object:

```{r}
(res_tm <- res_sp %>% 
  match_temporal(
    match_id = group,
    temporal_by = c("prcp" = "Water_course_level"),
    return_cubble = TRUE, 
    temporal_n_highest = 30,
    temporal_min_match = 15))
```

<!--We first pass in the two datasets, followed by the variables used for the temporal matching, specified in \code{temporal_by} (using the \code{by} syntax from \code{dplyr::join}). With \code{temporal_independent} we select precipitation (\code{prcp}) as the independent variable, and the goal is to see if the water level in the river reflects what is observed in precipitation (for nearby stations). Since we consider a full year of data, we set the number of peaks considered (\code{temporal_n_highest}) to 30 (slightly above the default value of 20), and \code{temporal_min_match} is raised accordingly.-->

This function returns a \code{cubble} object, with additional columns: \code{dist} storing the distance between matched stations, \code{group} summarizing spatial matching, and \code{n_match} showing the temporal matching.

Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b). The expected concurrent increase in precipitation and water level can be seen clearly.

```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 5, 6, 10) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))

res_tm_long <- res_tm %>%  
  face_temporal() %>%  
  unfold(group, type) %>%  
  rename(prcp = matched) %>%  
  group_by(group, type) %>%
  mutate(prcp = (prcp - min(prcp, na.rm = TRUE))/ (max(prcp, na.rm = TRUE) - min(prcp, na.rm = TRUE)))

vic_map <- ozmaps::abs_ste |> 
  filter(NAME == "Victoria") |> 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate_vic), 
             aes(x = long, y = lat, color = type), alpha = 0.2, fill = 0.2) +
  geom_point(data = res_tm, aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res_tm |> filter(type == "river"), aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

p2 <- res_tm_long %>%  
  ggplot(aes(x = date, y = prcp, group = type,color = type)) + 
  geom_line() + 
  facet_wrap(vars(group)) + 
  scale_color_brewer(palette = "Dark2", guide = "none") + 
  theme_bw() + 
  labs(x=  "date") + 
  scale_x_date(date_labels = "%b") + 
  labs(x = "Week", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom") 
```