### 2023-06-15 4-3-matching.Rmd

---
title: "4-3-water-level"
output: html_document
---

## River levels and rainfall in Victoria

Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Both `climate_vic` and `river` are cubble objects and we can get a summary of the 10 closest pairs:

```{r}
climate_vic <- climate_aus |>
  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
  mutate(type = "climate")
river <- cubble::river %>% mutate(type = "river") 
```

```{r echo = TRUE}
res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
print(res_sp, n = 20)
```

The result can also be returned as cubble objects with argument `return_cubble = TRUE`. The output is a list where each element is a paired cubble object and you may consider combining all the results into a single cubble with `bind_rows()`. Care needs to be taken on in the case when a site is close to two stations since by construction, cubble require unique rows in the nested form. From the summary table above, river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5). (Similarly river station `230200` is matched in group 7 and 8). One can either deselect one pair before binding the results, or take the list and work with the `purrr::map` syntax:

```{r echo = TRUE}
res_sp <- match_spatial(
  climate_vic, river, 
  spatial_n_group = 10, return_cubble = TRUE)
str(res_sp, max.level = 0)
(res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
```

For temporal matching, the variable water level (`Water_course_level`) from the river data will be matched to precipitation (`prcp`) in the weather station data. The variable identifying each matched group is `group` and the variable identifying the two datasets is `type`, and similarly, the cubble output can be returned with the argument `return_cubble = TRUE`. Here we select the four pairs with the highest number of matching peaks:

```{r echo = TRUE}
res_tm <- res_sp %>% 
  match_temporal(
    data_id = type, match_id = group,
    temporal_by = c("prcp" = "Water_course_level"),
    return_cubble = TRUE)
(res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
```

Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.

```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))

res_tm_long <- res_tm %>%  
  face_temporal() %>%  
  unfold(group, type) %>%  
  group_by(group, type) %>%
  mutate(matched = (matched - min(matched, na.rm = TRUE))/ 
           (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))

vic_map <- ozmaps::abs_ste |> 
  filter(NAME == "Victoria") |> 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate_vic), 
             aes(x = long, y = lat, color = type), 
             alpha = 0.2, fill = 0.2) +
  geom_point(data = res_tm %>% as_tibble(), 
             aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res_tm |> filter(type == "river") %>% as_tibble(), 
    aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

p2 <- res_tm_long %>%  
  ggplot(aes(x = date, y = matched, group = type,color = type)) + 
  geom_line() + 
  facet_wrap(vars(group)) + 
  scale_color_brewer(palette = "Dark2", guide = "none") + 
  theme_bw() + 
  labs(x=  "date") + 
  scale_x_date(date_labels = "%b") + 
  labs(x = "Week", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom") 
```