### 2023-06-18 3-advanced.Rmd

---
title: "3-advanced"
output: pdf_document
---

# Other features and considerations {#others}

## Data fusion and matching {#matching}

One common task when working with spatio-temporal data is to match nearby sites. For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources. In cubble, spatial and temporal matching are performed using the functions `match_spatial()` and `match_temporal()`. The `match_spatial()` function calculates the spatial distance between observations in two cubble objects. Various distance measures are available (check `sf::st_distance`). Analysts can specify the number of matched groups to output using the `spatial_n_group` argument (default to 4 groups) and the number of matches per group using the `spatial_n_group` argument (default to 1, one-to-one matching). The syntax to use `match_spatial()` is:

````
match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
````

The function `match_temporal()` calculates the similarity between time series within spatially matched groups. Two identifiers are required: one for separating each matched group (`match_id`) and one for separating the two data sources (`data_id`). The argument `temporal_by` uses the `by` syntax from dplyr's `*_join` to specify the temporal variables to match.

The similarity score between two time series is calculated using a matching function, which can be customised by the analysts. The matching function takes two time series as a list and returns a single numerical score. This allows for flexibility in using existing time series distance calculation implementation. By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window. The syntax to use `match_temporal()` is

````
match_temporal(
  <cubble_obj_from_match_spatial>, 
  data_id = , match_id = , 
  temporal_by = c("..." = "...")
)
````

## Interactive graphics

The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk]. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object. The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected. The long \code{cubble} will then highlight the corresponding series in the time series plot (c).

Linking is also available starting from the time series plot, by selecting points. This will activate rows having the same id in the long \code{cubble}. The corresponding rows in the nested \code{cubble} are activated, and highlighted on the map. (An illustration can be found in the appendix.) Note that this type of linking, both from the map or the time series, is what @CS07 would call categorical variable linking, where station id is the categorical variable.

```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots. The line plots and the map are constructed from shared \\code{crosstalk} objects (long and nested \\code{cubble}). When a station is selected on the map (a), the corresponding row in the nested \\code{cubble} will be activated. This will link to all the rows with the same id in the long \\code{cubble} (b) and update the line plot (c)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

## Spatio-temporal transformations {#st_transformation}

Sometimes, we wish to communicate spatial and temporal information collectively through visualisation. This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot. The concept of glyph maps was initially proposed in @Wickham2012-yr. The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.

The package `GGally` initially implement the glyph map. It uses `glyphs()` to calculate the axis transformation and then uses `geom_polygon()` to draw the map. In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation . The `geom_glyph()` requires four aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`. The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:

```
data |> 
  ggplot() +
  geom_glyph(aes(x_major = ..., x_minor = ..., 
                 y_major = ..., y_minor = ...))
```

Some useful controls over the glyph map includes:

  * polar coordinate glyph maps with \code{polar = TRUE}, 
  * adjust glyph size with arguments \code{width} and \code{height},
  * glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and 
  * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.