### 2023-06-18 4-3-matching.Rmd

---
title: "4-3-water-level"
output: html_document
---

## River levels and rainfall in Victoria

One common task when working with spatio-temporal data is to match nearby sites. For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources. In this example, we will introduce the spatial and temporal matching in cubble using an example on matching river level data with precipitation in Victoria, Australia.

The water level data collected by the Bureau of Meteorology, can be compared with the precipitation since rainfall can directly impact water level in river. Both `climate_vic` and `river` are cubble objects, and we can obtain a summary of the 10 closest pairs between them:

```{r}
climate_vic <- climate_aus |>
  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 76, 90)) |>
  mutate(type = "climate")
river <- cubble::river %>% mutate(type = "river") 
```

```{r echo = TRUE}
res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
print(res_sp, n = 20)
```

The result can also be returned as cubble objects by setting the argument `return_cubble = TRUE`. The output is be a list where each element is a paired cubble object. To combine all the results into a single cubble, you can use `bind_rows()`. In the case when a site in the second cubble (the `river` data here) is matched to two stations in the first cubble (`climate_vic` here), the binding may not be successful since cubble requires unique rows in the nested form. In the summary table above, the river station `226027` is matched to more than one weather station: `ASN00085072` (group 3) and `ASN00085298` (group 5). Similarly, the river station `230200` is matched in group 7 and 8). In such cases, you can either deselect one pair before combining, or work with the list output with the `purrr::map` syntax:

```{r echo = TRUE}
res_sp <- match_spatial(
  climate_vic, river, 
  spatial_n_group = 10, return_cubble = TRUE)
str(res_sp, max.level = 0)
(res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
```

For temporal matching, we match the variable `Water_course_level` from the river data to `prcp` in the weather station data. The variable `group` and `types` identify the matching group and the two datasets:

```{r echo = TRUE}
(res_tm <- res_sp %>% 
  match_temporal(
    data_id = type, match_id = group,
    temporal_by = c("prcp" = "Water_course_level")))
```

Similarly, the cubble output can be returned using the argument `return_cubble = TRUE`. Here we select the four pairs with the highest number of matching peaks:

```{r echo = TRUE}
res_tm <- res_sp %>% 
  match_temporal(
    data_id = type, match_id = group,
    temporal_by = c("prcp" = "Water_course_level"),
    return_cubble = TRUE)
(res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
```

Figure \ref{fig:matching} shows four matched pairs on the map (a) and standardized data as time series (b) with concurrent increasing.

```{r matching, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here."}
res_tm <- res_tm %>%  filter(group %in% c(1, 7, 6, 9))

res_tm_long <- res_tm %>%  
  face_temporal() %>%  
  unfold(group, type) %>%  
  group_by(group, type) %>%
  mutate(matched = (matched - min(matched, na.rm = TRUE))/ 
           (max(matched, na.rm = TRUE) - min(matched, na.rm = TRUE)))

vic_map <- ozmaps::abs_ste |> 
  filter(NAME == "Victoria") |> 
  rmapshaper::ms_simplify()

p1 <-ggplot() + 
  geom_sf(data = vic_map, fill = "grey95", color = "white") + 
  geom_point(data = dplyr::bind_rows(river, climate_vic), 
             aes(x = long, y = lat, color = type), 
             alpha = 0.2, fill = 0.2) +
  geom_point(data = res_tm %>% as_tibble(), 
             aes(x = long, y = lat, color = type)) +
  ggrepel::geom_label_repel(
    data = res_tm |> filter(type == "river") %>% as_tibble(), 
    aes(x = long, y = lat, label = group)) +
  scale_color_brewer(palette = "Dark2")  + 
  ggthemes::theme_map() +
  ggplot2::theme(legend.position = "bottom",
                 legend.text = element_text(size = 15),
                 legend.title = element_text(size = 15)) +
  ggplot2::labs(x = "Longitude", y = "Latitude") + 
  guides(color = guide_legend(override.aes = list(size=5)))

p2 <- res_tm_long %>%  
  ggplot(aes(x = date, y = matched, group = type,color = type)) + 
  geom_line() + 
  facet_wrap(vars(group)) + 
  scale_color_brewer(palette = "Dark2", guide = "none") + 
  theme_bw() + 
  labs(x=  "date") + 
  scale_x_date(date_labels = "%b") + 
  labs(x = "Week", y = "Precipitation/ water level")

(p1 | p2) + 
  patchwork::plot_layout(guides = "collect") + 
  plot_annotation(tag_levels = 'a')&
  theme(legend.position = "bottom") 
```