### 2023-06-18 4-5-interactive.Rmd

---
title: "4-5-interactive"
output: html_document
---

## Interactive graphics

Interactive graphics can be useful because they make it possible to look at the data in multiple ways on-the-fly. This is especially important for spatio-temporal data, where we would like to interactively connect spatial and temporal displays. This example describes the process of using the \pkg{cubble} package with the \pkg{crosstalk} package to build an interactive display connecting a map of Australia, with ribbon plots of temperature range observed at the stations. The purpose is to explore the variation of monthly temperature range over the country. Figure \ref{fig:interactive-linking} shows three snapshots of the interactivity.

The key steps are to convert both the nested and long forms of the data into shared \code{crosstalk} objects, and to plot these side-by-side. The two are linked by the station identifier.

``` r
clean <- climate_full |> ...

nested <- clean |> SharedData$new(~id, group = "cubble")
long <- face_temporal(clean) |> SharedData$new(~id, group = "cubble")

p1 <- nested |> ...
p2 <- long |> ...

crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)

```

Plot (a) shows the initial state of the interactive display: all locations are shown as dots on the map, coloured by temperature range, and the right plot shows the ribbons representing maximum to minimum for all stations. In plot (b) the "Mount Elizabeth" station, which shows a high variance colour on the initial map, is selected on the map and this produces the ribbon on the right. In plot (c) the lowest temperature in August is selected, which is "Thredbo" station on the left map. It was surprising to us that this was not a station in Tasmania, so for comparison a station in Tasmania is selected on the map to show in relation to Thredbo. We can see that Thredbo has a bigger winter dip in temperature, and although Tasmania is cold generally, it's temperatures are more constant

```{r eval = FALSE}
# This chunk generates the interactive graphic and save it as figures/linking.html, where the three screencasts in the figure comes from.
library("crosstalk")
library("plotly")

clean <- climate_full |>
  filter(!id %in% c("ASN00067033", "ASN00072091", "ASN00059040", 
                    "ASN00097053", "ASN00040856", "ASN00015548")) |>
  face_temporal() |> 
  mutate(month = lubridate::month(date)) |>
  group_by(month) |>
  summarise(
    tmax = mean(tmax, na.rm = TRUE),
    tmin = mean(tmin, na.rm = TRUE),
    diff = mean(tmax - tmin, na.rm = TRUE)
    ) |> 
  ungroup(month) |>
  mutate(dummy_date = as.Date(glue::glue("2021-{month}-01"))) |> 
  face_spatial() |> 
  mutate(temp_diff_var = ts |> pull(diff) |> var(na.rm = TRUE))

nested <- clean |> 
  SharedData$new(~id, group = "cubble")

long <- clean |> 
  face_temporal() |> 
  unfold(temp_diff_var) |> 
  arrange(temp_diff_var) |> 
  SharedData$new(~id, group = "cubble")

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)

p1 <- nested |> 
  ggplot() +
  geom_sf(data = state_map, fill = "grey95", color = "white") +
  geom_point(aes(x = long, y = lat,color = temp_diff_var, label = name)) +
  colorspace::scale_color_continuous_sequential(
    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
  labs(x = "Longitude", y = "Latitude") + 
  coord_sf() + 
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(),
    legend.position = "none"
    )

p2 <- long |> 
  ggplot(aes(x = dummy_date, group = id, 
             fill = temp_diff_var, color = temp_diff_var)) +
  geom_ribbon(aes(ymin = tmin, ymax = tmax), size = 0.1, alpha = 0.3) + 
  geom_line(aes(y = tmin), alpha = 0.1) + 
  geom_line(aes(y = tmax), alpha = 0.1) + 
  geom_point(aes(y = tmax), size = 0.1) + 
  geom_point(aes(y = tmin), size = 0.1) + 
  colorspace::scale_fill_continuous_sequential(
    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
  colorspace::scale_colour_continuous_sequential(
    "Rocket",  n_interp = 7, cmax = 90, rev = TRUE, 
    c2 = 40, l2= 85, c1 = 20, l1 = 30, name = "Var. temp. diff.") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") + 
  labs(x = "Month", y = "Temperature") + 
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(),
    legend.position = "bottom"
    )

out <- bscols(
  ggplotly(p1, width = 900, height = 700, tooltip = "label") |>
    highlight(
      on = "plotly_selected", 
      off = "plotly_deselect", 
      opacityDim = 0.1),
  ggplotly(p2, width = 1200, height = 700, tooltip = "label") |> 
    highlight(
      on = "plotly_selected", 
      off = "plotly_deselect", 
      opacityDim = 0.012),
  widths = c(5, 5)
)

htmltools::save_html(out, file = "figures/linking.html")

```

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row. The maximum and minimum temperatures are shown using a ribbon. The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo). Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))
```