### 2023-06-20 3-advanced.Rmd
nested/long all to spatial/temporal cubble

---
title: "3-advanced"
output: pdf_document
---

# Other features and considerations {#others}

## Data fusion and matching {#matching}

One common task when working with spatio-temporal data is to match nearby sites. For example, we may want to verify the location of an old list of stations with current stations, or we may want to match the data from different data sources. In cubble, spatial and temporal matching are performed using the functions `match_spatial()` and `match_temporal()`. The `match_spatial()` function calculates the spatial distance between observations in two cubble objects. Various distance measures are available (check `sf::st_distance`). Analysts can specify the number of matched groups to output using the `spatial_n_group` argument (default to 4 groups) and the number of matches per group using the `spatial_n_group` argument (default to 1, one-to-one matching). The syntax to use `match_spatial()` is:

````
match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
````

The function `match_temporal()` calculates the similarity between time series within spatially matched groups. Two identifiers are required: one for separating each matched group (`match_id`) and one for separating the two data sources (`data_id`). The argument `temporal_by` uses the `by` syntax from dplyr's `*_join` to specify the temporal variables to match.

The similarity score between two time series is calculated using a matching function, which can be customised by the analysts. The matching function takes two time series as a list and returns a single numerical score. This allows for flexibility in using existing time series distance calculation implementation. By default, cubble implements a simple peak matching algorithm (`match_peak`) that counts the number of peaks in two time series that fall within a specified temporal window. The syntax to use `match_temporal()` is

````
match_temporal(
  <cubble_obj_from_match_spatial>, 
  data_id = , match_id = , 
  temporal_by = c("..." = "...")
)
````

## Interactive graphics

The workflow with the \code{cubble} class works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk]. Figure \ref{fig:illu-interactive} illustrates the linking between a map and multiple time series in a \code{cubble} object. Both the map, generated from the spatial cubble, and the time series plot, generated from the temporal cubble, are shared crosstalk objects. When a user selects a location on the map (a), the corresponding site is highlighted. This selection activates a row in the spatial \code{cubble}, which is then communicated to the temporal \code{cubble}, resulting in the selection of all observations with the same ID in (b). Consequently the temporal \code{cubble} highlights the corresponding series in the time series plot (c). Linking can also be initiated from the time series plot, by selecting points on the time series. This action activates rows with the same ID in the temporal \code{cubble} and the row in the spatial \code{cubble}, which will be highlighted on the map. Please refer to the appendix for the corresponding illustration.

```{r illu-interactive, echo = FALSE, fig.align="center", out.height = "35%", out.width = "100%", fig.cap = "Linking between multiple plots. The line plots and the map are constructed from shared \\code{crosstalk} objects. When a station is selected on the map (a), the corresponding row in the spatial \\code{cubble} will be activated. This will link to all the rows with the same id in the temporal \\code{cubble} (b) and update the line plot (c)."}
knitr::include_graphics(here::here("figures/diagram-keynotes/diagram-keynotes.002.png"))
```

## Spatio-temporal transformations {#st_transformation}

Sometimes, we wish to communicate spatial and temporal information collectively through visualisation. This can be achieved through several graphical displays: one can make faceted maps across time, creating map animations, or constructing interactive graphics to link between map and time series plot. The concept of glyph maps was initially proposed in @Wickham2012-yr. The underlying idea is to transform the temporal coordinates into spatial coordinates so that time series plot can be displayed on the map.

The package `GGally` initially implement the glyph map. It uses `glyphs()` to calculate the axis transformation and then uses `geom_polygon()` to draw the map. In cubble, a ggproto implementation `geom_glyph()` performs the linear algebra internally as data transformation . The `geom_glyph()` requires four aesthetics: `x_major`, `y_major`, `x_minor`, and `y_minor`. The major axes are the outer spatial coordinates and the minor axes are the inner/ temporal coordinates:

```
data |> 
  ggplot() +
  geom_glyph(aes(x_major = ..., x_minor = ..., 
                 y_major = ..., y_minor = ...))
```

Some useful controls over the glyph map includes:

  * polar coordinate glyph maps with \code{polar = TRUE}, 
  * adjust glyph size with arguments \code{width} and \code{height},
  * glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE}) or free, and 
  * reference boxes and lines with \code{geom_glyph_box()} and \code{geom_glyph_line()}.