### 2023-06-21 main.tex

\documentclass[
]{jss}

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

\usepackage[utf8]{inputenc}

\author{
H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash
University \And Ursula Laa\\Universität für Bodenkultur Wiens
\AND Nicolas Langrené\\BNU-HKBU United International
College \And Patricia Menéndez\\Monash University
}
\title{\pkg{cubble}: An \proglang{R} Package for Organizing and
Wrangling Multivariate Spatio-temporal Data}

\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
Langrené, Patricia Menéndez}
\Plaintitle{cubble: An R Package for Organizing and Wrangling
Multivariate Spatio-temporal Data}

\Abstract{
Multivariate spatio-temporal data refers to multiple measurements taken
across space and time. For many analyses, spatial and time components
can be separately studied: for example, to explore the temporal trend of
one variable for a single spatial location, or to model the spatial
distribution of one variable at a given time. However for some studies,
it is important to analyze different aspects of the spatio-temporal data
simultaneously, like for instance, temporal trends of multiple variables
across locations. In order to facilitate the study of different portions
or combinations of spatio-temporal data, we introduce a new class,
\code{cubble}, with a suite of functions enabling easy slicing and
dicing on different spatio-temporal components. The proposed
\code{cubble} class ensures that all the components of the data are easy
to access and manipulate while providing flexibility for data analysis.
In addition, the \pkg{cubble} package facilitates visual and numerical
explorations of the data while easing data wrangling and modelling. The
\code{cubble} class and the functions provided in the \pkg{cubble}
\proglang{R} package equip users with the capability to handle
hierarchical spatial and temporal structures. The \code{cubble} class
and the tools implemented in the package are illustrated with different
examples of Australian climate data.
}

\Keywords{spatial, temporal, spatio
temporal, \proglang{R}, environmental data, exploratory data analysis}
\Plainkeywords{spatial, temporal, spatio temporal, R, environmental
data, exploratory data analysis}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    H. Sherry Zhang\\
    Monash University\\
    21 Chancellors Walk, Clayton VIC 3800 Australia\\
  E-mail: \email{huize.zhang@monash.edu}\\

      Dianne Cook\\
    Monash University\\
    21 Chancellors Walk, Clayton VIC 3800 Australia\\
  E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\

      Ursula Laa\\
    Universität für Bodenkultur Wiens\\
    Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
  E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\

      Nicolas Langrené\\
    BNU-HKBU United International College\\
    2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
  E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\

      Patricia Menéndez\\
    Monash University\\
    21 Chancellors Walk, Clayton VIC 3800 Australia\\
  E-mail: \href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}\\

  }

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{amsmath} \usepackage{array}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Spatio-temporal data has a spatial component referring to the location
of each observation and a temporal component that is recorded at regular
or irregular time intervals. It may also include multiple variables
measured at each spatial and temporal values. With spatio-temporal data,
one can fix the time to explore the spatial features of the data, fix
the spatial location/s to explore temporal aspects, or dynamically
explore the space and time simultaneously.

In order to computationally explore the spatial, temporal and
spatio-temporal faces of such data, the data needs to be stored and
represented under a specific data object that allows the user to query,
group and dissect all the data faces.

The Comprehensive \proglang{R} Archive Network (CRAN) task view
SpatioTemporal \citep{ctvspatiotemporal} gathers information about
\proglang{R} packages designed for spatio-temporal data and it has a
section on \emph{Representing data} that lists existing spatio-temporal
data representations used in \proglang{R}. Among them, \citet{spacetime}
summarises spatio-temporal data into three forms: time-wide, space-wide,
and long formats. The associated package \pkg{spacetime}
\citep{spacetime} implements four spatio-temporal layouts (full grid,
sparse grid, irregular, and trajectory) to handle different space and
time combinations. The package \pkg{stars} \citep{stars} has a new
implementation to use dense arrays to represent spatio-temporal cubes.
It also interfaces with the package \pkg{sf} \citep{sf}, commonly used
for wrangling spatial data, and the \pkg{tidyverse} \citep{tidyverse}
suite for general data wrangling and visualization in \proglang{R}.

Still, the data representation for spatio-temporal data can be further
extended and there are two reasons for this. Firstly, the raw data
sourced in the wild is less often presented in any one of the layouts
above, and fitting the raw data into a data object can sometimes be
difficult. More often, spatio-temporal data are collected in separate 2D
tables and analysts need to assemble them into a whole piece before
exploring the data. Examples of components of spatio-temporal data can
be 1) areal data recording the shape of a collection of areas of
interest; 2) geostatistical data storing the longitude and latitude
coordinates of locations, typically also with other metadata related to
the location, and; 3) temporal data of each location across time.

The other reason is about tidy data concepts \citep{tidydata} and how
they should be applied to spatio-temporal data. According to the tidy
data principles, data should be structured into 1) one row per
observation, 2) one column per variable, and 3) one type of data per
table. The long form data is preferred over wide data form given the
downstream packages such as \pkg{dplyr} \citep{dplyr} and \pkg{ggplot2}
\citep{ggplot2} for data wrangling and visualization. However, the long
form can be inefficient to store feature geometries, especially for
large multipolygons for hourly, daily or sub-daily periods over years,
which are extensively collected and handled, for example in time series
analysis. This poses the question of how to arrange spatial and temporal
variables in a way that would make data wrangling, visualizing and
analyzing spatio-temporal data easier.

This paper presents a new \proglang{R} package, \pkg{cubble}, which
addresses the two issues mentioned above. In the package, a new class,
also called cubble, is proposed to organize spatial and temporal
variables as two forms of a single data object so that they can be
wrangled separately or combined while being kept synchronized. Among the
four spacetime layouts in \citet{spacetime}, the \code{cubble} class can
be applied to full grid, sparse grid, or irregular, but not trajectory,
which is outside the scope of this work. The software is available from
the Comprehensive \proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/package=cubble}.

The rest of the paper is organized as follows: Section @ref(cubble)
presents the main design and functionality of the \pkg{cubble} package.
Section @ref(others) explains how the \pkg{cubble} package deals with
more advanced considerations, including data with hierarchical
structure, data matching and how the package fits with existing static
and interactive visualization tools. Moreover we also illustrate how the
\pkg{cubble} package deals with spatio-temporal data transformations.
Section @ref(examples) uses Australian weather station data and river
level data as examples to demonstrate the use of the package. An example
of how the \pkg{cubble} package handles Network Common Data Form
(NetCDF) data is also provided. Section @ref(conclude) discuss the paper
contributions and future directions.

\hypertarget{cubble}{%
\section{The cubble package}\label{cubble}}

\hypertarget{object}{%
\subsection{The cubble object}\label{object}}

Spatio-temporal data can encompass data with various spatial and
temporal characteristics and different data require different structures
for wrangling and analysis: climate weather stations record data at
fixed point location but may suffer from potential temporal data quality
issue such as missing data for certain days. GPS data tracks unique
point locations at different timestamps and is represented as
trajectories. Satellite imageries capture snapshots of landscapes at
selected time and is commonly structured as raster data. The
spatio-temporal data cubble address are those collected at unique fixed
locations while allowing for irregularity in the temporal dimension,
such as the weather station data. This corresponds to the full
space-time and sparse space-time layouts described in the spacetime
paper \citep{spacetime} and reproduced in Figure \ref{fig:layout}.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{/Users/hzha400/Documents/research/paper-cubble/figures/layout-1}

}

\caption[The full space-time and sparse space-time layouts handled by the cubble package]{The full space-time and sparse space-time layouts handled by the cubble package. Reproduced from Figure 1 in the spacetime paper \citep{spacetime}.}(\#fig:layout)
\end{figure}
\end{CodeChunk}

The cubble class is an S3 class built on tibble that allows the
spatio-temporal data to be wrangled in two forms: a nested form (spatial
cubble) and a long form (temporal cubble). It consists of two
subclasses:

\begin{itemize}
\tightlist
\item
  a spatial cubble with class
  \texttt{c("spatial\_cubble\_df",\ "cubble\_df")}
\item
  a emporal cubble with class
  \texttt{c("temporal\_cubble\_df",\ "cubble\_df")}
\end{itemize}

In a spatial cubble, spaital variables are organised as columns and
temporal variables are nested within a specialised \texttt{ts} column.
The cubble object printed below contains weather records in three
airport stations (the creation of a cubble object will be explained in
Section @ref(create)). This toy data is a subset of a larger data
\code{climate_aus} collected from Global Historical Climatology Network
Daily (GHCND). It records three airport stations located in Melbourne,
Australia and includes spatial variables such as station ID, longitude,
latitude, elevation, station name, World Meteorology Organisation ID.
The dataset contains temporal variables including precipitation, maximum
and minimum temperature, which can be read from the cubble header:

\begin{CodeChunk}
\begin{CodeInput}
R> cb_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form
# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id           long   lat  elev name              wmo_id ts               
  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
\end{CodeOutput}
\end{CodeChunk}

In a temporal cubble, temporal variables are expanded in the long form
and spatial variables are stored as a data attribute:

\begin{CodeChunk}
\begin{CodeInput}
R> cb_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, long form
# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00086038 2020-01-01     0  26.8  11  
2 ASN00086038 2020-01-02     0  26.3  12.2
3 ASN00086038 2020-01-03     0  34.5  12.7
4 ASN00086038 2020-01-04     0  29.3  18.8
5 ASN00086038 2020-01-05    18  16.1  12.5
# i 25 more rows
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{the-cubble-attributes}{%
\subsubsection{The cubble attributes}\label{the-cubble-attributes}}

A cubble object inherits the attributes from tibble (and its
subclasses): \texttt{class}, \texttt{row.names}, and \texttt{names}.
Additionally, it has three specialised attributes: \texttt{key},
\texttt{index}, and \texttt{coords}. Readers who are familiar with the
\texttt{key} and \texttt{index} attributes from the \texttt{tsibble}
package would already understand the two arguments. In cubble, the
\texttt{key} attribute identifies the row in the spatial cubble (given
the use of \texttt{tidyr::nest()} for nesting), and when combined with
the \texttt{index} argument, it identifies the row in the temporal
cubble. Currently, cubble only supports one variable as the key, and the
accepted temporal classes for index includes the base R classes
\texttt{Date}, \texttt{POSIXlt}, \texttt{POSIXct}, as well as tsibble's
\texttt{yearmonth}, \texttt{yearweek}, and \texttt{yearquarter} classes.

The \texttt{coords} attribute represents an ordered pair of coordinates.
It can be either an unprojected pair of longitude and latitude, or a
projected easting and northing value. The \texttt{sf} package is used
under the hood to calculate the bounding box, displayed in the header of
a spatial cubble.

The temporal cubble has a special attribute called \texttt{spatial} to
store the spatial variables. Shortcut functions are available to extract
attributes, for example, \code{spatial()} for extracting spatial
variables from the temporal cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> spatial(cb_long)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 3 x 6
  id           long   lat  elev name              wmo_id
  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl>
1 ASN00086038  145. -37.7  78.4 essendon airport   95866
2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870
3 ASN00086282  145. -37.7 113.  melbourne airport  94866
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{create}{%
\subsection{Creation and coercion}\label{create}}

In many cases, spatio-temporal data arrive in separate tables for
analysis. For example, in climate data, analysts may initially receive
station data containing geographic location information, recorded
variables and their recording periods. They can then query the temporal
variables using the stations of interest to obtain the relevant temporal
data. Alternatively, analyses may begin as purely spatial or temporal,
and analysts may obtain additional temporal or spatial data to expand
the result to spatio-temporal.

The function \code{make_cubble()} composes a \code{cubble} object from a
spatial table (\code{spatial}) and a temporal table (\code{temporal}),
along with three attributes introduced in the subsection @ref(object):
\code{key}, \code{index}, and \code{coords}. The following code creates
the spatial \code{cubble} object shown previously:

\begin{CodeChunk}
\begin{CodeInput}
R> make_cubble(spatial = stations, temporal = meteo,
+             key = id, index = date, coords = c(long, lat))
\end{CodeInput}
\end{CodeChunk}

Spatio-temporal data in other foreign objects can be coerced into a
\code{cubble} object with the function \code{as_cubble()}. This includes
casting from a \code{tibble} or \code{data.frame} with both spatial and
temporal information, a NetCDF object, a \code{stars} object
\citep{stars}, and a \code{sftime} object \citep{sftime}. The two
examples below show the casting from a tibble and a NetCDF object. The
dataset \code{climate_flat} combines the spatial data, \code{stations},
with the temporal data, \code{meteo}, into a single tibble object. It
can be coerced into a cubble using:

\begin{CodeChunk}
\begin{CodeInput}
R> climate_flat |> as_cubble(key = id, index = date, coords = c(long, lat))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form
# spatial:  [144.8321, -37.98, 145.0964, -37.6655], Missing CRS!
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id           long   lat  elev name              wmo_id ts               
  <chr>       <dbl> <dbl> <dbl> <chr>              <dbl> <list>           
1 ASN00086038  145. -37.7  78.4 essendon airport   95866 <tibble [10 x 4]>
2 ASN00086077  145. -38.0  12.1 moorabbin airport  94870 <tibble [10 x 4]>
3 ASN00086282  145. -37.7 113.  melbourne airport  94866 <tibble [10 x 4]>
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{functions-and-methods}{%
\subsection{Functions and methods}\label{functions-and-methods}}

Table @ref(tab:funs) summarises the functions implemented in the
\pkg{cubble} package and Table @ref(tab:methods) details the methods
implemented for each of the three cubble classes. The \code{cubble_df}
class handles methods that behave consistently in both spatial
(\code{spatial_cubble_df}) and temporal cubble
(\code{temporal_cubble_df}). When the method has distinct behavior for
the spatial cubble and temporal cubble, it is implemented separately for
each subclass.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1250}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8750}}@{}}
\caption{(\#tab:funs) Functions implemented in the \pkg{cubble}
package}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endhead
base R & \texttt{{[}}, \texttt{{[}{[}\textless{}-},
\texttt{names\textless{}-} \\
tidyverse & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify},
\texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate},
\texttt{arrange}, \texttt{filter}, \texttt{group\_by}, \texttt{ungroup},
\texttt{summarise}, \texttt{select}, \texttt{slice}, \texttt{rowwise},
\texttt{rename}, \texttt{bind\_rows}, \texttt{bind\_cols},
\texttt{relocate}, \texttt{type\_sum}, the slice family
(\texttt{slice\_head}, \texttt{slice\_tail}, \texttt{slice\_max},
\texttt{slice\_min}, \texttt{slice\_sample}) and the join family
(\texttt{left\_join}, \texttt{right\_join}, \texttt{inner\_join},
\texttt{full\_join}, \texttt{anti\_join}, \texttt{semi\_join}) \\
cubble & \texttt{as\_cubble}, \texttt{cubble}, \texttt{make\_cubble},
\texttt{check\_key} \texttt{face\_temporal}, \texttt{face\_spatial},
\texttt{unfold}, \texttt{key}, \texttt{key\_vars}, \texttt{key\_data},
\texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial},
\texttt{match\_sites}, \texttt{match\_spatial},
\texttt{match\_temporal}, \texttt{geom\_glyph},
\texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line},
\texttt{make\_spatial\_sf}, \texttt{make\_temporal\_tsibble},
\texttt{fill\_gaps}, and \texttt{scan\_gaps} \\
\bottomrule()
\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
\caption{(\#tab:methods) Methods implemented for the three \code{cubble}
classes.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Methods
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Methods
\end{minipage} \\
\midrule()
\endhead
\texttt{cubble\_df} &
\texttt{{[}{[}\textless{}-,\ dplyr\_col\_modify,\ key\_data,\ key\_vars,\ key,\ print} \\
\texttt{spatial\_cubble\_df} &
\texttt{{[},\ names\textless{}-,\ \ tbl\_sum,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ arrange,\ rename,\ rowwise,\ group\_by,\ ungroup,\ select,\ spatial,\ summarise,\ unfold,\ update\_cubble} \\
\texttt{temporal\_cubble\_df} &
\texttt{{[},\ names\textless{}-,\ tbl\_sum,\ arrange,\ dplyr\_reconstruct,\ dplyr\_row\_slice,\ face\_spatial,\ face\_temporal,\ unfold,\ fill\_gaps,\ group\_by,\ ungroup,\ \ rename,\ rowwise,\ scan\_gaps,\ select,\ spatial,\ summarise,\ tbl\_sum,\ bind\_rows,\ bind\_cols,\ update\_cubble} \\
\bottomrule()
\end{longtable}

Specifically, the pair of verbs, \code{face_temporal()} and
\code{face_spatial()}, pivot the cubble object between the spatial and
temporal face of the multivariate spatio-temporal cube, as illustrated
in Figure \ref{fig:face}. The code below uses the function
\code{face_temporal()} and \code{face_spatial()} to transform between a
spatial cubble (\code{cb_nested}) and a temporal one (\code{cb_long}),
introduced earlier in subsection @ref(object):

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_temporal(cb_nested), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_spatial(cb_long), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

The pair of verbs are exact inverse and apply both functions on a cubble
object will result in the object itself:

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_temporal(face_spatial(cb_long)), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/diagram-keynotes/diagram-keynotes.001}

}

\caption{An illustration of the function \code{face\_temporal()} and \code{face\_spatial()}: \code{face\_temporal()} converts a spatial cubble (neseted form) into a temporal cubble (long form) to focus on the temporal variables. Conversely, \code{face\_spatial()} transforms a temporal cubble into a spatial one to emphasize spatial variables.}(\#fig:face)
\end{figure}
\end{CodeChunk}

\hypertarget{compatibility-with-tsibble-and-sf}{%
\subsection{Compatibility with tsibble and
sf}\label{compatibility-with-tsibble-and-sf}}

Analysts often have their preferred spatial or temporal data structure
that they prefer to use for spatio-temporal analysis. For example, the
\code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is
commonly used in time series forecasting and similarly, the \code{sf}
class \citep{sf} is frequently used in spatial data science. In
\code{cubble}, analysts have the flexibility to combine these two
structures together by allowing the spatial component to also be an
\code{sf} object and the temporal component to also be a \code{tsibble}
object.

\hypertarget{using-a-tsibble-object-as-the-temporal-component}{%
\subsubsection{Using a tsibble object as the temporal
component}\label{using-a-tsibble-object-as-the-temporal-component}}

The \code{key} and \code{index} arguments in a \code{cubble} object
corresponds to the \code{tsibble} counterparts and they can be safely
omitted, if the temporal component is a \code{tsibble} object,
i.e.~\code{meteo_ts} in the example below. The \code{tsibble} class from
the input will be carried over to the \code{cubble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> ts_nested <- make_cubble(
+   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
R> (ts_long <- face_temporal(ts_nested))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, long form, [tsibble]
# temporal: 2020-01-01 -- 2020-01-10 [1D], no gaps
# spatial:  long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00086038 2020-01-01     0  26.8  11  
2 ASN00086038 2020-01-02     0  26.3  12.2
3 ASN00086038 2020-01-03     0  34.5  12.7
4 ASN00086038 2020-01-04     0  29.3  18.8
5 ASN00086038 2020-01-05    18  16.1  12.5
# i 25 more rows
\end{CodeOutput}
\begin{CodeInput}
R> class(ts_long)
\end{CodeInput}
\begin{CodeOutput}
[1] "temporal_cubble_df" "cubble_df"          "tbl_ts"            
[4] "tbl_df"             "tbl"                "data.frame"        
\end{CodeOutput}
\end{CodeChunk}

The temporal cubble shows \code{[tsibble]} in the header to indicate the
object also being in a \code{tbl_ts} class (\code{tsibble}). Methods
applies to the \code{tbl_ts} class can also be applied to the temporal
cubble objects, for example, checking whether the data contain temporal
gaps:

\begin{CodeChunk}
\begin{CodeInput}
R> ts_long %>% has_gaps()
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 3 x 2
  id          .gaps
  <chr>       <lgl>
1 ASN00086038 FALSE
2 ASN00086077 FALSE
3 ASN00086282 FALSE
\end{CodeOutput}
\end{CodeChunk}

An existing \code{cubble} object can promote its temporal component to a
\code{tsibble} object by applying \code{make_temporal_tsibble()}. The
following code illustrates this with the object \code{cb_long} created
in section @ref(create) and the promoted cubble object is equal to the
cubble object originally created from a \code{tsibble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> ts_long2 <- cb_long %>% make_temporal_tsibble() 
R> identical(ts_long2, ts_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{using-an-sf-object-as-the-spatial-component}{%
\subsubsection{Using an sf object as the spatial
component}\label{using-an-sf-object-as-the-spatial-component}}

Similarly, an \code{sf} object can be supplied as the spatial component
to create a \code{cubble} object, with the \code{coords} argument being
omitted. This opens up the possibility to represent fixed area with
polygons or multipolygons (see Applications @ref(covid)) and the
\code{coords} argument will be calculated as the centroids of the
(multi)polygons. The \code{[sf]} print in the cubble header suggest an
spatial component being also a \code{sf} object:

\begin{CodeChunk}
\begin{CodeInput}
R> (sf_nested <- make_cubble(
+   spatial = stations_sf, temporal = meteo, 
+   key = id, index = date))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form, [sf]
# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id           elev name   wmo_id  long   lat            geometry ts      
  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
1 ASN00086038  78.4 essen~  95866  145. -37.7 (144.9066 -37.7276) <tibble>
2 ASN00086077  12.1 moora~  94870  145. -38.0   (145.0964 -37.98) <tibble>
3 ASN00086282 113.  melbo~  94866  145. -37.7 (144.8321 -37.6655) <tibble>
\end{CodeOutput}
\begin{CodeInput}
R> class(sf_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] "spatial_cubble_df" "cubble_df"         "sf"               
[4] "tbl_df"            "tbl"               "data.frame"       
\end{CodeOutput}
\end{CodeChunk}

The \code{sf} functions applicable to a \code{cubble} object have been
listed in Table @ref(tab:funs) and the following code shows how to
perform coordinate transformation with \code{st_transform} on a cubble
object:

\begin{CodeChunk}
\begin{CodeInput}
R> sf_nested %>% sf::st_transform(crs = "EPSG:3857")
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form, [sf]
# spatial:  [16122635.6225205, -4576600.8687746, 16152057.3639371,
#   -4532279.35567565], WGS 84
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id           elev name   wmo_id  long   lat            geometry ts      
  <chr>       <dbl> <chr>   <dbl> <dbl> <dbl>         <POINT [°]> <list>  
1 ASN00086038  78.4 essen~  95866  145. -37.7 (16130929 -4541016) <tibble>
2 ASN00086077  12.1 moora~  94870  145. -38.0 (16152057 -4576601) <tibble>
3 ASN00086282 113.  melbo~  94866  145. -37.7 (16122636 -4532279) <tibble>
\end{CodeOutput}
\end{CodeChunk}

The counterpart to promote the spatial component in an existing
\code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:

\begin{CodeChunk}
\begin{CodeInput}
R> cb_nested %>% make_spatial_sf() 
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form, [sf]
# spatial:  [144.8321, -37.98, 145.0964, -37.6655], WGS 84
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id           long   lat  elev name   wmo_id ts                  geometry
  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
1 ASN00086038  145. -37.7  78.4 essen~  95866 <tibble> (144.9066 -37.7276)
2 ASN00086077  145. -38.0  12.1 moora~  94870 <tibble>   (145.0964 -37.98)
3 ASN00086282  145. -37.7 113.  melbo~  94866 <tibble> (144.8321 -37.6655)
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{tidyverse}{%
\subsection{Comparison to other spatio-temporal
classes}\label{tidyverse}}

In \proglang{R}, there are other existing spatio-temporal data structure
and this section compares and contrasts \pkg{cubble} with other existing
alternative, specifically \pkg{stars} and \pkg{sftime}. The \pkg{stars}
package \citep{stars} uses an array structure, as oppose to tibble, to
represent multivariate spatio-temporal data. While both \pkg{stars} and
\pkg{cubble} support vector and raster data, it is a matter of choice on
which structure to use given the application. Analysts working on
satellite imageries may prefer the array structure in \pkg{stars}, while
others originally working with spatio-temporal data in 2D data frames
may find \pkg{cubble} easier to adopt from their existing computing
workflow.

The \pkg{sftime} package \citep{sftime} also builds from a tibble object
and its focus is on handling irregular spatio-temporal data. This means
\pkg{sftime} can also handle full space-time grids and sparse space-time
layouts represented in \pkg{cubble}, but \pkg{cubble} uses nesting to
avoid storing spatial variables repetitively at each timestamp. This
provides memory efficiency when data is observed frequent, i.e.~daily or
sub-daily, or the spatial geometry is expensive to repeat, i.e.~polygons
or multipolygons. Consider the \code{climate_aus} data in the
\pkg{cubble} package with 639 stations observed daily in a single year
2020. The created \code{sftime} object is approximately 14 times larger
than the corresponding \code{cubble} object (118.24 MB vs.~8.52 MB).

\hypertarget{others}{%
\section{Other features and considerations}\label{others}}

\hypertarget{matching}{%
\subsection{Data fusion and matching}\label{matching}}

One common task when working with spatio-temporal data is to match
nearby sites. For example, we may want to verify the location of an old
list of stations with current stations, or we may want to match the data
from different data sources. In cubble, spatial and temporal matching
are performed using the functions \texttt{match\_spatial()} and
\texttt{match\_temporal()}. The \texttt{match\_spatial()} function
calculates the spatial distance between observations in two cubble
objects. Various distance measures are available (check
\texttt{sf::st\_distance}). Analysts can specify the number of matched
groups to output using the \texttt{spatial\_n\_group} argument (default
to 4 groups) and the number of matches per group using the
\texttt{spatial\_n\_group} argument (default to 1, one-to-one matching).
The syntax to use \texttt{match\_spatial()} is:

\begin{verbatim}
match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
\end{verbatim}

The function \texttt{match\_temporal()} calculates the similarity
between time series within spatially matched groups. Two identifiers are
required: one for separating each matched group (\texttt{match\_id}) and
one for separating the two data sources (\texttt{data\_id}). The
argument \texttt{temporal\_by} uses the \texttt{by} syntax from dplyr's
\texttt{*\_join} to specify the temporal variables to match.

The similarity score between two time series is calculated using a
matching function, which can be customised by the analysts. The matching
function takes two time series as a list and returns a single numerical
score. This allows for flexibility in using existing time series
distance calculation implementation. By default, cubble implements a
simple peak matching algorithm (\texttt{match\_peak}) that counts the
number of peaks in two time series that fall within a specified temporal
window. The syntax to use \texttt{match\_temporal()} is

\begin{verbatim}
match_temporal(
  <cubble_obj_from_match_spatial>, 
  data_id = , match_id = , 
  temporal_by = c("..." = "...")
)
\end{verbatim}

\hypertarget{interactive-graphics}{%
\subsection{Interactive graphics}\label{interactive-graphics}}

The workflow with the \code{cubble} class works well with an interactive
graphics pipeline (e.g., \citet{buja1988elements};
\citet{buja1996interactive}; \citet{sutherland2000orca};
\citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in
\proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}. Figure
\ref{fig:illu-interactive} illustrates the linking between a map and
multiple time series in a \code{cubble} object. Both the map, generated
from the spatial cubble, and the time series plot, generated from the
temporal cubble, are shared crosstalk objects. When a user selects a
location on the map (a), the corresponding site is highlighted. This
selection activates a row in the spatial \code{cubble}, which is then
communicated to the temporal \code{cubble}, resulting in the selection
of all observations with the same ID in (b). Consequently the temporal
\code{cubble} highlights the corresponding series in the time series
plot (c). Linking can also be initiated from the time series plot, by
selecting points on the time series. This action activates rows with the
same ID in the temporal \code{cubble} and the row in the spatial
\code{cubble}, which will be highlighted on the map. Please refer to the
appendix for the corresponding illustration.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth,height=0.35\textheight]{figures/diagram-keynotes/diagram-keynotes.002}

}

\caption{Linking between multiple plots. The line plots and the map are constructed from shared \code{crosstalk} objects. When a station is selected on the map (a), the corresponding row in the spatial \code{cubble} will be activated. This will link to all the rows with the same id in the temporal \code{cubble} (b) and update the line plot (c).}(\#fig:illu-interactive)
\end{figure}
\end{CodeChunk}

\hypertarget{st_transformation}{%
\subsection{Spatio-temporal transformations}\label{st_transformation}}

Sometimes, we wish to communicate spatial and temporal information
collectively through visualisation. This can be achieved through several
graphical displays: one can make faceted maps across time, creating map
animations, or constructing interactive graphics to link between map and
time series plot. The concept of glyph maps was initially proposed in
\citet{Wickham2012-yr}. The underlying idea is to transform the temporal
coordinates into spatial coordinates so that time series plot can be
displayed on the map.

The package \texttt{GGally} initially implement the glyph map. It uses
\texttt{glyphs()} to calculate the axis transformation and then uses
\texttt{geom\_polygon()} to draw the map. In cubble, a ggproto
implementation \texttt{geom\_glyph()} performs the linear algebra
internally as data transformation . The \texttt{geom\_glyph()} requires
four aesthetics: \texttt{x\_major}, \texttt{y\_major},
\texttt{x\_minor}, and \texttt{y\_minor}. The major axes are the outer
spatial coordinates and the minor axes are the inner/ temporal
coordinates:

\begin{verbatim}
data |> 
  ggplot() +
  geom_glyph(aes(x_major = ..., x_minor = ..., 
                 y_major = ..., y_minor = ...))
\end{verbatim}

Some useful controls over the glyph map includes:

\begin{itemize}
\tightlist
\item
  polar coordinate glyph maps with \code{polar = TRUE},
\item
  adjust glyph size with arguments \code{width} and \code{height},
\item
  glyph scale to fixed (\code{global_rescale} defaults to \code{TRUE})
  or free, and
\item
  reference boxes and lines with \code{geom_glyph_box()} and
  \code{geom_glyph_line()}.
\end{itemize}

\hypertarget{examples}{%
\section{Applications}\label{examples}}

Five examples are chosen to illustrate different aspects of the
\pkg{cubble} package: creating a \code{cubble} object from two
Coronavirus (COVID) data tables with the complication of differing
location names, using spatial transformations to make a glyph map of
seasonal temperature changes, matching river level data and weather
station records for analysis of water supply, reading NetCDF format data
to reproduce a climate reanalysis plot, and the workflow to create
complex interactive linked plots.

\hypertarget{covid}{%
\subsection{Victoria COVID spatio-temporal incidence and
spread}\label{covid}}

Since the start of the pandemic, the Victoria State Government in
Australia has been providing daily COVID counts per Logal Government
Area (LGA). This data can be combined with map polygon data, available
from the Australian Bureau of Statistics (ABS), to visualize COVID
incidence and spread. In the \pkg{cubble} package, the COVID count data
(\code{covid}) and the LGA information (\code{lga}) are available as a
\code{tsibble} object and an \code{sf} object respectively.

A cubble object can be created from separate spatial and temporal
component using the \code{make_cubble()} function, introduced in Section
@ref(create). The \code{by} argument is used to specify the joining
variable from the two component using the \code{*_join()} \code{by}
syntax:

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- make_cubble(lga, covid, by = c("lga_name_2018" = "lga"))
\end{CodeInput}
\begin{CodeOutput}
Warning: st_centroid assumes attributes are constant over geometries
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Use `check_key()` to check on the unmatched key
The cubble is created only with sites having both spatial and
temporal information
\end{CodeOutput}
\end{CodeChunk}

The warning message suggests the slight difference of LGA encoding used
by Victoria government and ABS and prompts analysts to use the function
\code{check_key()} to identify the mismatches:

\begin{CodeChunk}
\begin{CodeInput}
R> (check_res <- check_key(
+   spatial = lga, temporal = covid, 
+   by = c("lga_name_2018" = "lga")
+ ))
\end{CodeInput}
\begin{CodeOutput}
$paired
# A tibble: 78 x 2
  spatial        temporal      
  <chr>          <chr>         
1 Alpine (S)     Alpine (S)    
2 Ararat (RC)    Ararat (RC)   
3 Ballarat (C)   Ballarat (C)  
4 Banyule (C)    Banyule (C)   
5 Bass Coast (S) Bass Coast (S)
# i 73 more rows

$potential_pairs
# A tibble: 2 x 2
  spatial             temporal    
  <chr>               <chr>       
1 Kingston (C) (Vic.) Kingston (C)
2 Latrobe (C) (Vic.)  Latrobe (C)

$others
$others$spatial
[1] "No usual address (Vic.)"               
[2] "Migratory - Offshore - Shipping (Vic.)"

$others$temporal
[1] "Interstate" "Overseas"   "Unknown"   
\end{CodeOutput}
\end{CodeChunk}

The result of the \code{check_key()} function is a list containing three
elements: 1) matched keys from both tables, 2) potentially pairable
keys, and 3) others. Here, the main mismatch arises from the two LGAs:
Kingston and Latrobe (Kingston is a LGA in both Victoria and South
Australia and Latrobe is a LGA in both Victoria and Tasmania). Analysts
can modify the input spatial and temporal data accordingly and recreate
the cubble object:

\begin{CodeChunk}
\begin{CodeInput}
R> lga2 <- lga |>
+   rename(lga = lga_name_2018) |> 
+   mutate(lga = ifelse(lga == "Kingston (C) (Vic.)", "Kingston (C)", lga),
+          lga = ifelse(lga == "Latrobe (C) (Vic.)", "Latrobe (C)", lga)) |>
+   filter(!lga %in% check_res$others$spatial) 
R>   
R> covid2 <- covid |> filter(!lga %in% check_res$others$temporal)
R> 
R> (cb <- make_cubble(spatial = lga2, temporal = covid2))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: lga [80], index: date, nested form, [sf]
# spatial:  [140.961682, -39.1591895, 149.976291, -33.9804256], WGS 84
# temporal: date [date], n [dbl], avg_7day [dbl]
  lga             long   lat                                   geometry ts      
  <chr>          <dbl> <dbl>                         <MULTIPOLYGON [°]> <list>  
1 Alpine (S)      147. -36.9 (((146.7258 -36.45922, 146.8033 -36.45139~ <tbl_ts>
2 Ararat (RC)     143. -37.5 (((143.1807 -37.73152, 143.1793 -37.73242~ <tbl_ts>
3 Ballarat (C)    144. -37.5 (((143.6622 -37.57241, 143.6686 -37.53844~ <tbl_ts>
4 Banyule (C)     145. -37.7 (((145.1357 -37.74091, 145.1331 -37.74281~ <tbl_ts>
5 Bass Coast (S)  146. -38.5 (((145.5207 -38.30667, 145.5418 -38.30938~ <tbl_ts>
# i 75 more rows
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{historicaltmax}{%
\subsection{Australian historical maximum
temperature}\label{historicaltmax}}

The GHCN provides daily climate measures from stations across the world.
The data used here (\code{historical_tmax}) is a subset extracted using
the package \pkg{rnoaa} \citep{rnoaa}, containing the records of maximum
temperature for 237 Australian stations from \ensuremath{\infty{}}
through \ensuremath{-\infty{}} and provides information also on the
latitude, longitude and elevation of each of the stations. The goal of
this example is to compare the monthly average maximum temperature
between two periods, 1971-1975 and 2016-2020, for stations in Victoria
and New South Wales (NSW), using a glyph map.

First, the stations need to be filtered to those in Victoria and NSW by
using the station identifiers, stored within the 11 digits of the
\code{id} variable entries. The country code is in the first 5 digits
(Australia is represented by ``ASN00'') and the next 6 digits encode the
station following the Australian Bureau of Meteorology (BOM) \citep{bom}
coding protocols. The NSW stations correspond to entries in the range
46-75 and the Victorian stations to 76-90. Filtering Victoria and NSW
stations is a \emph{spatial operation} and hence uses the spatial
\code{cubble}:

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- historical_tmax |>
+   filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
\end{CodeInput}
\end{CodeChunk}

Next, the monthly maximum average temperature is calculated for both
periods. This is a \emph{temporal operation} requiring a switch into the
temporal \code{cubble} using the \code{face_temporal()} function. In
addition, a new indicator for the two time periods of interest is
created before the calculation of monthly averages:

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |>
+   face_temporal() |> 
+   group_by(
+     yearmonth = tsibble::make_yearmonth(
+       year = ifelse(lubridate::year(date) > 2015, 2016, 1971),
+       month = lubridate::month(date))
+       )|>
+   summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
+   mutate(group = as.factor(lubridate::year(yearmonth)),
+          month = lubridate::month(yearmonth))
\end{CodeInput}
\end{CodeChunk}

A quick check on the number of observations for each location is made,
revealing that there are several with less than 24 observations -- these
stations lack temperature values for some months. In this example, those
stations are removed by switching to the spatial \code{cubble} to
operate on the spatial component over time, and then, move back into the
temporal \code{cubble} (to make the glyph map):

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |> 
+   face_spatial() |> 
+   rowwise() %>% 
+   filter(nrow(ts) == 24) |>
+   face_temporal()
\end{CodeInput}
\end{CodeChunk}

In order to create a glyph map displaying the monthly series (Figure
\ref{fig:basic-manip}), the spatial variables need to be unfolded with
the temporal variables. The reason being that the major (\code{long},
\code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be
on the same table to create the glyph map. The \code{geom_glyph()}
function does both the transformation and the plotting.

\begin{CodeChunk}
\begin{CodeInput}
R> ggplot() +
+   geom_sf(data = MAP_DATA, ...) +
+   geom_glyph(data = tmax,
+              aes(x_major = long, x_minor = month,
+                  y_major = lat, y_minor = tmax,
+                  group = interaction(id, group), color = group),
+              width = 1, height = 0.5) + 
+   ...
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/basic-manip-1}

}

\caption[A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020)]{A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020). The corresponding average time series for the cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW.}(\#fig:basic-manip)
\end{figure}
\end{CodeChunk}

Glyph maps work well to explore temporal patterns across spatial
locations, particularly when the spatial locations are gridded. In this
example, they are irregularly spaced, which can result in overlapping
glyphs obscuring each other.

\hypertarget{river-levels-and-rainfall-in-victoria}{%
\subsection{River levels and rainfall in
Victoria}\label{river-levels-and-rainfall-in-victoria}}

One common task when working with spatio-temporal data is to match
nearby sites. For example, we may want to verify the location of an old
list of stations with current stations, or we may want to match the data
from different data sources. In this example, we will introduce the
spatial and temporal matching in cubble using an example on matching
river level data with precipitation in Victoria, Australia.

The water level data collected by the Bureau of Meteorology, can be
compared with the precipitation since rainfall can directly impact water
level in river. Both \texttt{climate\_vic} and \texttt{river} are cubble
objects, and we can obtain a summary of the 10 closest pairs between
them:

\begin{CodeChunk}
\begin{CodeInput}
R> res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
R> print(res_sp, n = 20)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 10 x 4
   from        to      dist group
   <chr>       <chr>    [m] <int>
 1 ASN00088051 406213 1838.     1
 2 ASN00084145 222201 2185.     2
 3 ASN00085072 226027 3282.     3
 4 ASN00080015 406704 4034.     4
 5 ASN00085298 226027 4207.     5
 6 ASN00082042 405234 6153.     6
 7 ASN00086038 230200 6167.     7
 8 ASN00086282 230200 6928.     8
 9 ASN00085279 224217 7431.     9
10 ASN00080091 406756 7460.    10
\end{CodeOutput}
\end{CodeChunk}

The result can also be returned as cubble objects by setting the
argument \texttt{return\_cubble\ =\ TRUE}. The output is be a list where
each element is a paired cubble object. To combine all the results into
a single cubble, you can use \texttt{bind\_rows()}. In the case when a
site in the second cubble (the \texttt{river} data here) is matched to
two stations in the first cubble (\texttt{climate\_vic} here), the
binding may not be successful since cubble requires unique rows in the
nested form. In the summary table above, the river station
\texttt{226027} is matched to more than one weather station:
\texttt{ASN00085072} (group 3) and \texttt{ASN00085298} (group 5).
Similarly, the river station \texttt{230200} is matched in group 7 and
8). In such cases, you can either deselect one pair before combining, or
work with the list output with the \texttt{purrr::map} syntax:

\begin{CodeChunk}
\begin{CodeInput}
R> res_sp <- match_spatial(
+   climate_vic, river, 
+   spatial_n_group = 10, return_cubble = TRUE)
R> str(res_sp, max.level = 0)
\end{CodeInput}
\begin{CodeOutput}
List of 10
\end{CodeOutput}
\begin{CodeInput}
R> (res_sp <- res_sp[-c(5, 8)] %>% bind_rows())
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [16], index: date, nested form, [sf]
# spatial:  [144.5203, -38.144913, 148.4667, -36.128657], WGS 84
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id      long   lat  elev name  wmo_id ts       type              geometry
  <chr>  <dbl> <dbl> <dbl> <chr>  <dbl> <list>   <chr>          <POINT [°]>
1 ASN00~  145. -37.0 290   rede~  94859 <tibble> clim~  (144.5203 -37.0194)
2 406213  145. -37.0  NA   CAMP~     NA <tibble> river (144.5403 -37.01512)
3 ASN00~  148. -37.7  62.7 orbo~  95918 <tibble> clim~  (148.4667 -37.6922)
4 222201  148. -37.7  NA   SNOW~     NA <tibble> river  (148.451 -37.70739)
5 ASN00~  147. -38.1   4.6 east~  94907 <tibble> clim~  (147.1322 -38.1156)
# i 11 more rows
# i 2 more variables: group <int>, dist [m]
\end{CodeOutput}
\end{CodeChunk}

For temporal matching, we match the variable
\texttt{Water\_course\_level} from the river data to \texttt{prcp} in
the weather station data. The variable \texttt{group} and \texttt{types}
identify the matching group and the two datasets:

\begin{CodeChunk}
\begin{CodeInput}
R> (res_tm <- res_sp %>% 
+   match_temporal(
+     data_id = type, match_id = group,
+     temporal_by = c("prcp" = "Water_course_level")))
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 8 x 2
  group match_res
  <int>     <dbl>
1     1        30
2     2         5
3     3        14
4     4        20
5     6        23
# i 3 more rows
\end{CodeOutput}
\end{CodeChunk}

Similarly, the cubble output can be returned using the argument
\texttt{return\_cubble\ =\ TRUE}. Here we select the four pairs with the
highest number of matching peaks:

\begin{CodeChunk}
\begin{CodeInput}
R> res_tm <- res_sp %>% 
+   match_temporal(
+     data_id = type, match_id = group,
+     temporal_by = c("prcp" = "Water_course_level"),
+     return_cubble = TRUE)
R> (res_tm <- res_tm %>% bind_rows() %>% filter(group %in% c(1, 7, 6, 9)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [8], index: date, nested form, [sf]
# spatial:  [144.5203, -37.8817, 147.572223, -36.8472], WGS 84
# temporal: date [date], matched [dbl]
  id         long   lat  elev name  wmo_id type              geometry group
  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <chr>          <POINT [°]> <int>
1 ASN00088~  145. -37.0 290   rede~  94859 clim~  (144.5203 -37.0194)     1
2 406213     145. -37.0  NA   CAMP~     NA river (144.5403 -37.01512)     1
3 ASN00082~  146. -36.8 502   stra~  95843 clim~  (145.7308 -36.8472)     6
4 405234     146. -36.9  NA   SEVE~     NA river (145.6828 -36.88701)     6
5 ASN00086~  145. -37.7  78.4 esse~  95866 clim~  (144.9066 -37.7276)     7
# i 3 more rows
# i 3 more variables: dist [m], ts <list>, match_res <dbl>
\end{CodeOutput}
\end{CodeChunk}

Figure \ref{fig:matching} shows four matched pairs on the map (a) and
standardized data as time series (b) with concurrent increasing.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/matching-1}

}

\caption[Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b)]{Weather stations and river gauges with matched pairs labelled on the map (a) and plotted across time (b). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation. The numbers (1, 7, 6, 9) indicate the group index derived from spatial matching, only those that were selectd by temporal matching are shown here.}(\#fig:matching)
\end{figure}
\end{CodeChunk}

\hypertarget{era5-climate-reanalysis-data}{%
\subsection{ERA5: climate reanalysis
data}\label{era5-climate-reanalysis-data}}

The ERA5 reanalysis \citep{hersbach2020era5} provides hourly estimates
of atmospheric, land and oceanic climate variables on a global scale.
The hourly pressure level data in the NetCDF format can be downloaded
from Copernicus Climate Data Store (CDS) \citep{cds} or via the
\pkg{ecmwfr} package \citep{ecwmfr}. This example reproduces the row
created from ERA5 reanalysis data shown in Figure 19 by
\citet{hersbach2020era5}. The plot shows the southern polar vortex
splitting into two on 2002-09-26, and further splitting into four on
2002-10-04. Further explanation of why this is interesting can be found
in the figure source, and also in \citet{simmons2020global} and
\citet{simmons2005ecmwf}.

To work with NetCDF data in R, several packages are available, including
\code{ncdf4}, \code{RNetCDF}, and \code{tidync}. The following code
converts a NetCDF object of class \code{ncdf4} into a cubble object:

\begin{CodeChunk}
\begin{CodeInput}
R> raw <- ncdf4::nc_open(here::here("data/era5-pressure.nc"))
\end{CodeInput}
\end{CodeChunk}

Analysts can extract a subset of the NetCDF data with the arguments
\code{vars}, \code{long_range} and \code{lat_range}. In this example,
the variables q (specific humidity) and z (geopotential) are selected
and the coordinates are subsetted to every degree in longitude and
latitude:

\begin{CodeChunk}
\begin{CodeInput}
R> (dt <- as_cubble(raw, vars = c("q", "z"),
+                  long_range = seq(-180, 180, 1), lat_range = seq(-88, -15, 1)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [26640], index: time, nested form
# spatial:  [-180, -88, 179, -15], Missing CRS!
# temporal: time [date], q [dbl], z [dbl]
     id  long   lat ts              
  <int> <dbl> <dbl> <list>          
1     1  -180   -15 <tibble [8 x 3]>
2     2  -179   -15 <tibble [8 x 3]>
3     3  -178   -15 <tibble [8 x 3]>
4     4  -177   -15 <tibble [8 x 3]>
5     5  -176   -15 <tibble [8 x 3]>
# i 26,635 more rows
\end{CodeOutput}
\end{CodeChunk}

Once the NetCDF data is coerced into a cubble object, subsequent
analysis can be conducted to filter on the date of interest, scale the
variable specific humidity and create visualisation in ggplot as in
Figure \ref{fig:netcdf}.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/research/paper-cubble/figures/netcdf-1}

}

\caption[A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002]{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020) to illustrate the break-up of sourthern polar vortex in late September and early October 2002. The polar vortex, signalled by the high specific humidity, splits into two on 2002-09-26 and further splits into four on 2002-10-04.}(\#fig:netcdf)
\end{figure}
\end{CodeChunk}

\hypertarget{australian-temperature-range}{%
\subsection{Australian temperature
range}\label{australian-temperature-range}}

Interactive graphics can be useful because they make it possible to look
at the data in multiple ways on-the-fly. This is especially important
for spatio-temporal data, where we would like to interactively connect
spatial and temporal displays. This example describes the process of
using the \pkg{cubble} package with the \pkg{crosstalk} package to build
an interactive display connecting a map of Australia, with ribbon plots
of temperature range observed at the stations. The purpose is to explore
the variation of monthly temperature range over the country. Figure
\ref{fig:interactive-linking} shows three snapshots of the
interactivity.

The key steps are to convert both the nested and long forms of the data
into shared \code{crosstalk} objects, and to plot these side-by-side.
The two are linked by the station identifier.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
\end{Highlighting}
\end{Shaded}

Plot (a) shows the initial state of the interactive display: all
locations are shown as dots on the map, coloured by temperature range,
and the right plot shows the ribbons representing maximum to minimum for
all stations. In plot (b) the ``Mount Elizabeth'' station, which shows a
high variance colour on the initial map, is selected on the map and this
produces the ribbon on the right. In plot (c) the lowest temperature in
August is selected, which is ``Thredbo'' station on the left map. It was
surprising to us that this was not a station in Tasmania, so for
comparison a station in Tasmania is selected on the map to show in
relation to Thredbo. We can see that Thredbo has a bigger winter dip in
temperature, and although Tasmania is cold generally, it's temperatures
are more constant

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-north} \includegraphics[width=1\linewidth,height=0.23\textheight]{figures/linking-lower}

}

\caption[Exploring temperature variation using linking of a map and seasonal display]{Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a particular location on the map (here Mount Elizabeth) produces the plot in the second row. The maximum and minimum temperatures are shown using a ribbon. The bottom row first selects the lowest temperature in August in the seasonal display, which highlights the corresponding station on the map (Thredbo). Another  station, located in the Tasmania Island, is then selected to compare its temperature variation with the Thredbo station.}(\#fig:interactive-linking)
\end{figure}
\end{CodeChunk}

\hypertarget{conclude}{%
\section{Conclusion}\label{conclude}}

This paper presents an \proglang{R} package \pkg{cubble} for organizing,
manipulating and visualizing spatio-temporal data. The package
introduces a new data class for spatio-temporal data, \code{cubble},
that connects the time invariant and varying variables and that allows
the user to work with a nested and long form of the data. This work adds
capabilities into the spatio-temporal practitioners toolbox to integrate
it with a tidy data framework. The data structure and functions
introduced in this package can be used and combined with existing
spatial data analysis packages such as \pkg{sf}, data wrangling packages
such as \pkg{dplyr}, and visualization packages such as \pkg{ggplot2},
\pkg{plotly} and \pkg{leaflet}.

Numerous examples are provided in the main text, appendix and package
vignettes. These include creating and coercing data with mismatched
names, handling hierarchical data, matching time series spatially and
temporally, reproducing ERA5 plots from NetCDF data. Visualization of
the \code{cubble} objects can be done with interactive graphic pipelines
using \pkg{crosstalk}, \pkg{plotly} and \pkg{leaflet}.

There are several possible future directions for the work. The data
structure only described fixed spatial sites, and it could be useful to
provide tools to accommodate moving coordinates as might be encountered
in animal movement data. That could be achieved with a list-column for
the location coordinates, and an additional form that these locations
can be pivoted into, like the long form for temporal variables. For
multiple measured variables, the \pkg{cubble} package could be
integrated with dimension reduction methods, and dynamic graphics for
multiple dimensions such as a tour \citep{wickham_tourr_2011}.

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

This work is funded by a Commonwealth Scientific and Industrial Research
Organisation (CSIRO) Data61 Scholarship and started while Nicolas
Langrené was affiliated with CSIRO's Data61. The article is created
using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown}
\citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article}
template. The source code for reproducing this paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-cubble}.

\bibliography{references.bib}

\end{document}